---
output:
  pdf_document: default
  html_document: default
title: "Homogeneous Chain Binomial Epidemic Inference"
author: "Benjamen Simon"
date: "September 2022"
knit: (function(inputFile, encoding) { 
          rmarkdown::render(
            inputFile,
            encoding = encoding, 
            output_dir = "../Reports") 
            })
header-includes:
  - \usepackage{amsmath}
  - \usepackage{enumerate}
  - \usepackage{bm}
  - \usepackage[T1]{fontenc}
---

```{r setup, include=FALSE}

# Rmarkdown settings
knitr::opts_chunk$set(echo = FALSE, cache = TRUE, eval= TRUE, message = FALSE, cache.path = "Caches/", fig.path = "Figures/cbHomo/")

# Default to non-scientific notation
options(scipen=999)

# Suppress summarise info
options(dplyr.summarise.inform = FALSE)

```

```{r Packages}

  ## Data manipulation
  library(tidyverse)

  ## Plotting
  library(ggplot2)
  library(ggcorrplot)
  source("VisualisationFunctions.R")
  library(plotMCMC) # autocorrelation plots and other diagnostic plots

  ## MCMC Diagnostics
  library(coda)

  ## Epidemics
  library(EpidemicR) # For simulating and making inference on GSE epidemics

```

# Load in the epidemic

The epidemic we have simulated has a modal number of infected individuals as roughly 25% of the population. We ran the inference in Julia.

```{r Load epidemic}

epi_list <- readRDS(file="../Data/cbHomo/epi_list.RData")

# Associated infection rate matrix
  params <- epi_list[[1]]
  
# Simulated states
  sim_states <- epi_list[[2]]
  
# Simulated events
  sim_events <- epi_list[[3]]

# Inference (I known)
  res_dI_known <- read.csv(file="../Results/cbHomo/res_dI_known.csv")
  other_res_dI_known <- read.csv(file="../Results/cbHomo/other_res_dI_known.csv")
  
# Inference (I unknown)
  res_dI_unknown <- read.csv(file="../Results/cbHomo/res.csv")
  other_res_dI_unknown <- read.csv(file="../Results/cbHomo/other_res.csv")
  
# Priors
  # d_β_core = Uniform(0, 10)
  # d_δ_core = Uniform(0, 10)

```




# Making inference (I known)

```{r Summarise the results 1}

## Summarise the results

summary(res_dI_known[-(1:5000), 1:2])

summary(other_res_dI_known[-(1:5000), c(1,5)])

```

```{r}

hex_beta_delta_I_known = gg_hex_plot_cbhomo(res_dI_known[-c(1:10000), ], params_true, c(0.24,0.25,  0.24,0.22,  0.24,0.04,  0.24,0.01), c(0, 0.275), c(0, 0.26))

hex_beta_delta_I_known

ggsave(filename = "../Visualisation/cbHomo/hex_beta_delta_I_known.png", plot = hex_beta_delta_I_known, width = 10, height = 8)

```

```{r Plot the results 1}

## Plot the results

trace_plot_1 = gg_trace_plot_cbhomo(res_dI_known, params, 10000, c(40000, 0.025, 40000, 0.025), c(0, 0.2, 0, 0.2))

hist_plot_1 = gg_hist_plot_cbhomo(res_dI_known, params, 10000, c(0.04, 10, 0.04, 11), c(0, 0.25, 0, 0.25))

hist_plot_w_prior_1 = gg_hist_plot_w_prior_cbhomo(res_dI_known, params, 10000, c(0, 10, 0, 10), c(0.04, 10, 0.04, 11), c(0, 0.25, 0, 0.25))

## Save the results

ggsave(filename = "../Visualisation/cbHomo/trace_I_known.png", plot = trace_plot_1, width = 10, height = 8)

ggsave(filename = "../Visualisation/cbHomo/hist_I_known.png", plot = hist_plot_1, width = 10, height = 8)

ggsave(filename = "../Visualisation/cbHomo/hist_w_prior_I_known.png", plot = hist_plot_w_prior_1, width = 10, height = 8)


```

The MCMC algorithm for all parameters took `r round(end.time - start.time, 2)` minutes to run.

The algorithm estimates the mean, median, and quartiles of $\beta_1$ to be `r mean(res_one[-(1:5000),2])` (`r quantile(res_one[-(1:5000),2], probs = c(.25, .5, .75))`) compared to the true value of $\beta_1 =$  `r params[1]`, with an effective sample size of `r round(coda::effectiveSize(res_one[,2]))`. 

The algorithm estimates the mean, median, and quartiles of $\beta_2$ to be `r mean(res_one[-(1:5000),3])` (`r quantile(res_one[-(1:5000),3], probs = c(.25, .5, .75))`) compared to the true value of $\beta =$  `r params[2]`, with an effective sample size of `r round(coda::effectiveSize(res_one[,3]))`. 

It estimates the mean, median, and quartiles of $\delta$ to be `r mean(res_one[-(1:5000),3])` (`r quantile(res_one[-(1:5000),3], probs = c(.25, .5, .75))`) compared to the true values of $\delta =$ `r params[3]`, with an effective sample size of `r round(coda::effectiveSize(res_one[,3]))`.

It estimates the mean, median, and quartiles of $d$ to be `r mean(res_one[-(1:5000),4])` (`r quantile(res_one[-(1:5000),4], probs = c(.25, .5, .75))`) compared to the true values of $d =$ `r params[4]`, with an effective sample size of `r round(coda::effectiveSize(res_one[,4]))`.


```{r}

png("../Visualisation/cbHomo/autocorr_Iknown.png", width=1000, height=800)
plotAuto(res_dI_known[, c(1,2)], lag.max=50)
dev.off()

```


# Making inference (I unknown)


```{r Summarise the results 2}

## Summarise the results

summary(res_dI_unknown[-(1:5000), 1:2])

summary(other_res_dI_unknown[-(1:5000), c(1,5)])

```

```{r}

hex_beta_delta_I_unknown = gg_hex_plot_cbhomo(res_dI_unknown[-c(1:10000), ], params_true, c(0.03,0.25,  0.03,0.22,  0.24,0.04,  0.24,0.01), c(0, 0.275), c(0, 0.26))

hex_beta_delta_I_unknown

ggsave(filename = "../Visualisation/cbHomo/hex_beta_delta_I_unknown.png", plot = hex_beta_delta_I_unknown, width = 10, height = 8)

```




```{r}

contour_I_unknown = gg_contour_plot_cbhomo(res_dI_unknown, params, 10000, c(0.025,0.175), c(0.025,0.175))

contour_I_unknown

ggsave(filename = "../Visualisation/cbHomo/contour_I_unknown.png", plot = contour_I_unknown, width = 10, height = 8)

```





```{r Plot the results 2}

## Plot the results

trace_plot_2 = gg_trace_plot_cbhomo(res_dI_unknown, params, 10000, c(40000, 0.02, 40000, 0.02), c(0, 0.2, 0, 0.2))

hist_plot_2 = gg_hist_plot_cbhomo(res_dI_unknown, params, 10000, c(0.2, 10, 0.2, 11), c(0, 0.25, 0, 0.25))

hist_plot_w_prior_2 = gg_hist_plot_w_prior_cbhomo(res_dI_unknown, params, 10000, c(0, 10, 0, 10), c(0.2, 10, 0.2, 11), c(0, 0.25, 0, 0.25))

## Save the results

ggsave(filename = "../Visualisation/cbHomo/trace_I_unknown.png", plot = trace_plot_2, width = 10, height = 8)

ggsave(filename = "../Visualisation/cbHomo/hist_I_unknown.png", plot = hist_plot_2, width = 10, height = 8)

ggsave(filename = "../Visualisation/cbHomo/hist_w_prior_I_unknown.png", plot = hist_plot_w_prior_2, width = 10, height = 8)


```

The MCMC algorithm for all parameters took `r round(end.time - start.time, 2)` minutes to run.

The algorithm estimates the mean, median, and quartiles of $\beta_1$ to be `r mean(res.reparam_one[-(1:5000),2])` (`r quantile(res.reparam_one[-(1:5000),2], probs = c(.25, .5, .75))`) compared to the true value of $\beta_1 =$  `r params[1]`, with an effective sample size of `r round(coda::effectiveSize(res.reparam_one[,2]))`. 

The algorithm estimates the mean, median, and quartiles of $\beta_2$ to be `r mean(res.reparam_one[-(1:5000),3])` (`r quantile(res.reparam_one[-(1:5000),3], probs = c(.25, .5, .75))`) compared to the true value of $\beta =$  `r params[2]`, with an effective sample size of `r round(coda::effectiveSize(res.reparam_one[,3]))`. 

It estimates the mean, median, and quartiles of $\delta$ to be `r mean(res.reparam_one[-(1:5000),3])` (`r quantile(res.reparam_one[-(1:5000),3], probs = c(.25, .5, .75))`) compared to the true values of $\delta =$ `r params[3]`, with an effective sample size of `r round(coda::effectiveSize(res.reparam_one[,3]))`.

It estimates the mean, median, and quartiles of $d$ to be `r mean(res.reparam_one[-(1:5000),4])` (`r quantile(res.reparam_one[-(1:5000),4], probs = c(.25, .5, .75))`) compared to the true values of $d =$ `r params[4]`, with an effective sample size of `r round(coda::effectiveSize(res.reparam_one[,4]))`.


```{r}

png("../Visualisation/cbHomo/autocorr_Iunknown.png", width=1000, height=800)
plotAuto(res_dI_unknown[, c(1,2)], lag.max=50)
dev.off()

```
