---
output:
  pdf_document: default
  html_document: default
title: "Unpaired Homogeneous General Stochastic Epidemic Inference"
author: "Benjamen Simon"
date: "September 2022"
knit: (function(inputFile, encoding) { 
          rmarkdown::render(
            inputFile,
            encoding = encoding, 
            output_dir = "../Reports") 
            })
header-includes:
  - \usepackage{amsmath}
  - \usepackage{enumerate}
  - \usepackage{bm}
  - \usepackage[T1]{fontenc}
---

```{r setup, include=FALSE}

# Rmarkdown settings
knitr::opts_chunk$set(echo = FALSE, cache = TRUE, eval= TRUE, message = FALSE, cache.path = "Caches/", fig.path = "Figures/HeteroUnif/")

# Default to non-scientific notation
options(scipen=999)

# Suppress summarise info
options(dplyr.summarise.inform = FALSE)

```

```{r Packages}

  ## Data manipulation
  library(tidyverse)

  ## Plotting
  library(ggplot2)
  library(ggcorrplot)
  source("VisualisationFunctions.R")

  ## MCMC Diagnostics
  library(coda)

  ## Epidemics
  library(EpidemicR) # For simulating and making inference on GSE epidemics

  ## Write up
  library(xtable)

```

# Load in the epidemic

The epidemic we have simulated has a modal number of infected individuals as roughly 25% of the population

```{r Load epidemic}

epi_list <- readRDS(file="../Data/HeteroUnif/epi_list.RData")

# Distance matrix
  spatial <- epi_list[[1]]
  xy.coords <- spatial[[1]]
  distance_mat <- spatial[[2]]

# Associated infection rate matrix
  params <- epi_list[[2]]
  rate_mat <- epi_list[[3]]
   
# Simulated epidemic
  Homo_sim <- epi_list[[4]]

# Total number of infected
  n_I = sum(Homo_sim[,2] < Inf)
  
# Who is infected
  inf.ids <- which(Homo_sim[,2] < Inf)

# Infection times
  inf.times <- Homo_sim[,2]

# Removal times
  rem.times <- Homo_sim[,3]
  
# Plot the epidemic
  epi_plot = EpidemicR::Plot_epidemic(xy.coords, Homo_sim, init.inf = c(1))

```




# Making inference (Base)

We now make inference treating $\beta_1$ and $\beta_2$ as independent parameters.

```{r Performing analysis 1}

set.seed(1105)

## Run the algorithm

start.time <- Sys.time()

res.list_one <- MCMC(N.its = 100000, N = 100, inf.ids, rem.times, distance_mat, 
                     lambda.b1 = 0.001, nu.b1 = 1, lambda.b2 = 0.001, nu.b2 = 1, lambda.g = 0.001 , nu.g = 1,
                     inc.beta1 = list(NA, T), inc.beta2 = list(NA, T), inc.dist = list(NA, T),
                     inc.inf.times = list(NA, T), inc.gamma = list(NA, T),
                     d.upper = 15, sigmab1 = 1, sigmab2 = 5, sigmad = 0.04, infupdate = 5)

res_one <- res.list_one[[1]]

end.time <- Sys.time()

```

```{r Summarise the results 1}

## Summarise the results

summary(res_one[-(1:5000), 2:5])

Avg_acc_probs(res.list_one, F)

```



```{r Plot the results 1}

## Plot the results

trace_plot_betas_1 = gg_trace_plot_betas(res_one, params, 10000, c(20000, 0.025, 18000, 0.0013), c(0, 0.03, 0, 0.0015))

trace_plot_gd_1 = gg_trace_plot_gd(res_one, params, 10000, c(3000, 0.025, 17000, 3), c(0.01, 0.25, 2.5, 7))

# aaa = gg_trace_plot_bp(res_one, params, 10000, c(20000, 0.025, 18000, 0.25), c(0, 0.03, 0, 0.3))

hist_plot_1 = gg_hist_plot(res_one, params, 10000, c(0.02, 90, 0.001, 1500, 0.2, 8, 8, 0.4), c(0, 0.03, 0, 0.0017, 0, 0.35, 2, 12))

## Save the results

ggsave(filename = "../Visualisation/HeteroUnif/trace_betas.png", plot = trace_plot_betas_1, width = 10, height = 8)

ggsave(filename = "../Visualisation/HeteroUnif/trace_gd.png", plot = trace_plot_gd_1, width = 10, height = 8)

ggsave(filename = "../Visualisation/HeteroUnif/hist_plot.png", plot = hist_plot_1, width = 10, height = 8)


```

The MCMC algorithm for all parameters took `r round(end.time - start.time, 2)` minutes to run.

The algorithm estimates the mean, median, and quartiles of $\beta_1$ to be `r mean(res_one[-(1:5000),2])` (`r quantile(res_one[-(1:5000),2], probs = c(.25, .5, .75))`) compared to the true value of $\beta_1 =$  `r params[1]`, with an effective sample size of `r round(coda::effectiveSize(res_one[,2]))`. 

The algorithm estimates the mean, median, and quartiles of $\beta_2$ to be `r mean(res_one[-(1:5000),3])` (`r quantile(res_one[-(1:5000),3], probs = c(.25, .5, .75))`) compared to the true value of $\beta =$  `r params[2]`, with an effective sample size of `r round(coda::effectiveSize(res_one[,3]))`. 

It estimates the mean, median, and quartiles of $\gamma$ to be `r mean(res_one[-(1:5000),3])` (`r quantile(res_one[-(1:5000),3], probs = c(.25, .5, .75))`) compared to the true values of $\gamma =$ `r params[3]`, with an effective sample size of `r round(coda::effectiveSize(res_one[,3]))`.

It estimates the mean, median, and quartiles of $d$ to be `r mean(res_one[-(1:5000),4])` (`r quantile(res_one[-(1:5000),4], probs = c(.25, .5, .75))`) compared to the true values of $d =$ `r params[4]`, with an effective sample size of `r round(coda::effectiveSize(res_one[,4]))`.



# Making inference (Reparameterisation)

We now rerun the analysis under a reparameterised model in order to see if this has any effect on our ability to make efficient inference.

```{r Performing analysis reparam}

set.seed(1106)

## Run the algorithm

start.time <- Sys.time()

res.list.reparam_one <- MCMC_reparam(N.its = 100000, N = 100, inf.ids, rem.times, distance_mat, 
                     lambda.b1 = 0.001, nu.b1 = 1, lambda.g = 0.001 , nu.g = 1,
                     inc.beta1 = list(NA, T), inc.p = list(NA, T), inc.dist = list(NA, T),
                     inc.inf.times = list(NA, T), inc.gamma = list(NA, T),
                     d.upper = 15, sigmap = 0.01, sigmad = 0.1, infupdate = 2)

res.reparam_one <- res.list.reparam_one[[1]]

end.time <- Sys.time()

```

```{r Summarise the results reparam 1}

## Summarise the results

summary(res.reparam_one[-(1:5000), 2:5])

Avg_acc_probs(res.list.reparam_one, T)

```



```{r Plot the results reparam 1}

## Plot the results

trace_plot_betas_reparam_1 = gg_trace_plot_betas(res.reparam_one, params, 10000, c(23000, 0.025, 23000, 0.0013), c(0, 0.03, 0, 0.0015))

trace_plot_gd_reparam_1 = gg_trace_plot_gd(res.reparam_one, params, 10000, c(3000, 0.025, 20000, 3), c(0.01, 0.25, 2.5, 7))

trace_plot_bp_reparam_1 = gg_trace_plot_bp(res.reparam_one, params, 10000, c(23000, 0.025, 27000, 0.175), c(0, 0.03, 0, 0.2))

hist_plot_betas_reparam_1 = gg_hist_plot(res.reparam_one, params, 10000, c(0.0175, 90, 0.001, 1500, 0.2, 8, 8, 0.4), c(0, 0.0225, 0, 0.00125, 0, 0.25, 2, 10))

hist_plot_bp_reparam_1 = gg_hist_plot_reparam(res.reparam_one, params, 10000, c(0.0175, 100, 0.1, 10, 0.2, 8, 8, 0.4), c(0, 0.0225, 0, 0.15, 0, 0.25, 2, 10))

## Save the results

ggsave(filename = "../Visualisation/HeteroUnif/trace_betas_reparam.png", plot = trace_plot_betas_reparam_1, width = 10, height = 8)

ggsave(filename = "../Visualisation/HeteroUnif/trace_gd_reparam.png", plot = trace_plot_gd_reparam_1, width = 10, height = 8)

ggsave(filename = "../Visualisation/HeteroUnif/trace_bp_reparam.png", plot = trace_plot_bp_reparam_1, width = 10, height = 8)

ggsave(filename = "../Visualisation/HeteroUnif/hist_plot_betas_reparam.png", plot = hist_plot_betas_reparam_1, width = 10, height = 8)

ggsave(filename = "../Visualisation/HeteroUnif/hist_plot_bp_reparam.png", plot = hist_plot_bp_reparam_1, width = 10, height = 8)

```

The MCMC algorithm for all parameters took `r round(end.time - start.time, 2)` minutes to run.

The algorithm estimates the mean, median, and quartiles of $\beta_1$ to be `r mean(res.reparam_one[-(1:5000),2])` (`r quantile(res.reparam_one[-(1:5000),2], probs = c(.25, .5, .75))`) compared to the true value of $\beta_1 =$  `r params[1]`, with an effective sample size of `r round(coda::effectiveSize(res.reparam_one[,2]))`. 

The algorithm estimates the mean, median, and quartiles of $\beta_2$ to be `r mean(res.reparam_one[-(1:5000),3])` (`r quantile(res.reparam_one[-(1:5000),3], probs = c(.25, .5, .75))`) compared to the true value of $\beta =$  `r params[2]`, with an effective sample size of `r round(coda::effectiveSize(res.reparam_one[,3]))`. 

It estimates the mean, median, and quartiles of $\gamma$ to be `r mean(res.reparam_one[-(1:5000),3])` (`r quantile(res.reparam_one[-(1:5000),3], probs = c(.25, .5, .75))`) compared to the true values of $\gamma =$ `r params[3]`, with an effective sample size of `r round(coda::effectiveSize(res.reparam_one[,3]))`.

It estimates the mean, median, and quartiles of $d$ to be `r mean(res.reparam_one[-(1:5000),4])` (`r quantile(res.reparam_one[-(1:5000),4], probs = c(.25, .5, .75))`) compared to the true values of $d =$ `r params[4]`, with an effective sample size of `r round(coda::effectiveSize(res.reparam_one[,4]))`.



