---
output:
  pdf_document: default
  html_document: default
title: "Heterogeneous General Stochastic Epidemic Inference"
author: "Benjamen Simon"
date: "September 2022"
knit: (function(inputFile, encoding) { 
          rmarkdown::render(
            inputFile,
            encoding = encoding, 
            output_dir = "../Reports") 
            })
header-includes:
  - \usepackage{amsmath}
  - \usepackage{enumerate}
  - \usepackage{bm}
  - \usepackage[T1]{fontenc}
---

```{r setup, include=FALSE}

# Rmarkdown settings
knitr::opts_chunk$set(echo = FALSE, cache = TRUE, eval= TRUE, message = FALSE, cache.path = "Caches/", fig.path = "Figures/HeteroUnif/")

# Default to non-scientific notation
options(scipen=999)

# Suppress summarise info
options(dplyr.summarise.inform = FALSE)

```

```{r Packages}

  ## Data manipulation
  library(tidyverse)

  ## Plotting
  library(ggplot2)
  library(ggcorrplot)
  source("VisualisationFunctions.R")
  library(plotMCMC) # autocorrelation plots and other diagnostic plots


  ## MCMC Diagnostics
  library(coda)

  ## Epidemics
  library(EpidemicR) # For simulating and making inference on GSE epidemics

  ## Write up
  library(xtable)

```

# Load in the epidemic

The epidemic we have simulated has a modal number of infected individuals as roughly 25% of the population

```{r Load epidemic}

epi_list <- readRDS(file="../Data/HeteroUnif/epi_list.RData")

# Distance matrix
  spatial <- epi_list[[1]]
  xy.coords <- spatial[[1]]
  distance_mat <- spatial[[2]]

# Associated infection rate matrix
  params <- epi_list[[2]]
  rate_mat <- epi_list[[3]]
   
# Simulated epidemic
  Homo_sim <- epi_list[[4]]

# Total number of infected
  n_I = sum(Homo_sim[,2] < Inf)
  
# Who is infected
  inf.ids <- which(Homo_sim[,2] < Inf)

# Infection times
  inf.times <- Homo_sim[,2]

# Removal times
  rem.times <- Homo_sim[,3]
  
# Plot the epidemic
  epi_plot = EpidemicR::Plot_epidemic(xy.coords, Homo_sim, init.inf = c(1))
  
  epi_curve_plot = epidemic_curves_plot(inf.times, rem.times, inf.ids)
  epi_curve_plot

```
```{r}

  png("simulateddataplane.png", width = 6, height = 6, units = "in", res = 300)
  EpidemicR::Plot_epidemic(xy.coords, Homo_sim, init.inf = c(1))
  dev.off()
  
```

```{r}
# Step 2: Define distance thresholds
thresholds <- c(1, 3, 5, 7, 10, 12, 15, 20)

# Step 3: Function to calculate proportion within each threshold
calculate_proportion <- function(threshold, distance_mat) {
  # Exclude diagonal and upper triangle (duplicates) using lower.tri
  within_threshold <- sum(distance_mat[lower.tri(distance_mat)] <= threshold)
  total_pairs <- sum(lower.tri(distance_mat))

  proportion_within_threshold <- within_threshold / total_pairs *100
  return(proportion_within_threshold)
}

# Step 4: Apply the function for each threshold
proportions <- sapply(thresholds, function(x) calculate_proportion(x, distance_mat))

# Step 5: Display or use the results
result <- data.frame(Threshold = thresholds, Proportion = proportions)
print(result)
```


```{r}
# Step 2: Define distance thresholds
thresholds <- c(1, 3, 5, 7, 10, 12, 15, 20)

# Step 3: Function to calculate proportion within each threshold
calculate_proportion <- function(threshold, distance_mat, inf.ids) {
  # Exclude diagonal and upper triangle (duplicates) using lower.tri
  marked_distances <- distance_mat[inf.ids, inf.ids]
  within_threshold <- sum(marked_distances[lower.tri(marked_distances)] <= threshold)
  total_pairs <- sum(lower.tri(marked_distances))
  
  

  proportion_within_threshold <- within_threshold / total_pairs *100
  return(proportion_within_threshold)
}

# Step 4: Apply the function for each threshold
proportions <- sapply(thresholds, function(x) calculate_proportion(x, distance_mat, inf.ids))

# Step 5: Display or use the results
result <- data.frame(Threshold = thresholds, Proportion = proportions)
print(result)

```








# Making inference (Base)

We now make inference treating $\beta_1$ and $\beta_2$ as independent parameters.

```{r Performing analysis 1}

set.seed(1105)

## Run the algorithm

start.time <- Sys.time()

res.list_one <- MCMC(N.its = 100000, N = 100, inf.ids, rem.times, distance_mat, 
                     lambda.b1 = 1, nu.b1 = 1, lambda.b2 = 1, nu.b2 = 1, lambda.g = 1 , nu.g = 1,
                     inc.beta1 = list(0.007, T), inc.beta2 = list(0.00007, T), inc.dist = list(5, T),
                     inc.inf.times = list(NA, T), inc.gamma = list(0.11, T),
                     d.upper = 15, sigmab1 = 0.5, sigmab2 = 1.2, sigmad = 0.2, infupdate = 2)

res_one <- res.list_one[[1]]

end.time <- Sys.time()

print(c("Total Run Time:", end.time - start.time))

```
```{r}
coda::effectiveSize(res_one)

apply(res_one, 2, function(column) {
  quantiles <- quantile(column, c(0.025, 0.975))
  rounded_quantiles <- signif(quantiles, digits = 3)  # Round to 3 significant figures
  return(rounded_quantiles)
})


apply(res_one[ , ], 2, function(column) {
  std_dev <- sd(column)  # Calculate standard deviation
  rounded_std_dev <- signif(std_dev, digits = 3)  # Round to 3 significant figures
  return(rounded_std_dev)
})

```




```{r Summarise the results 1}

## Summarise the results

summary(res_one[-(1:5000), 2:5])

Avg_acc_probs(res.list_one, F)

```


```{r}

hex_b1_gamma_1 = gg_hex_plot_beta1_gamma(res_one[-(1:5000),], params_true, c(0.08,0.25,  0.08,0.22,  0.08,0.04,  0.08,0.01), c(0, 0.091), c(0, 0.26))

hex_b2_gamma_1 = gg_hex_plot_beta2_gamma(res_one[-(1:5000),], params_true, c(0.0019,0.25,  0.0019,0.22,  0.0019,0.04,  0.0019,0.01), c(0, 0.0022), c(0, 0.26))

hex_b1_b2_1 = gg_hex_plot_beta1_beta2(res_one[-(1:5000),], params_true, c(0.036,0.0017,  0.075,0.0017,  0.035,0.0019,  0.075,0.0019), c(0, 0.091), c(0, 0.0019))

hex_b1_d_1 = gg_hex_plot_beta1_d(res_one[-(1:5000),], params_true, c(0.036,7.5,  0.075,7.5,  0.035,6.5,  0.075,6.5), c(0, 0.091), c(1.35, 8))

hex_b2_d_1 = gg_hex_plot_beta2_d(res_one[-(1:5000),], params_true, c(0.00165,8,  0.00165,7.25,  0.0008,8,  0.0008,7.25), c(0, 0.0019), c(1.35, 8))

hex_d_gamma_1 = gg_hex_plot_d_gamma(res_one[-(1:5000),], params_true, c(1.7,0,  3.2,0,  5,0,  7,0), c(1.35, 8), c(0, 0.26))


ggsave(filename = "../Visualisation/HeteroUnif/hex_b1_gamma.png", plot = hex_b1_gamma_1, width = 10, height = 8)
ggsave(filename = "../Visualisation/HeteroUnif/hex_b2_gamma.png", plot = hex_b2_gamma_1, width = 10, height = 8)
ggsave(filename = "../Visualisation/HeteroUnif/hex_b1_b2.png", plot = hex_b1_b2_1, width = 10, height = 8)
ggsave(filename = "../Visualisation/HeteroUnif/hex_b1_d.png", plot = hex_b1_d_1, width = 10, height = 8)
ggsave(filename = "../Visualisation/HeteroUnif/hex_b2_d.png", plot = hex_b2_d_1, width = 10, height = 8)
ggsave(filename = "../Visualisation/HeteroUnif/hex_d_gamma.png", plot = hex_d_gamma_1, width = 10, height = 8)

```



```{r}

contour_b1_gamma_1 = gg_contour_plot_b1_gamma(res_one, params, 10000, c(0, 0.04), c(0, 0.175))

contour_b2_gamma_1 = gg_contour_plot_b2_gamma(res_one, params, 10000, c(0, 0.0022), c(0, 0.175))

contour_b1_b2_1 = gg_contour_plot_b1_b2(res_one, params, 10000, c(0, 0.015), c(0, 0.0019))

contour_b1_d_1 = gg_contour_plot_b1_d(res_one, params, 10000, c(0, 0.015), c(3, 6.25))

contour_b2_d_1 = gg_contour_plot_b2_d(res_one, params, 10000, c(0, 0.0019), c(1.35, 6.25))

contour_d_gamma_1 = gg_contour_plot_d_gamma(res_one, params, 10000, c(1.35, 6.25), c(0, 0.26))


ggsave(filename = "../Visualisation/HeteroUnif/contour_b1_gamma.png", plot = contour_b1_gamma_1, width = 10, height = 8)
ggsave(filename = "../Visualisation/HeteroUnif/contour_b2_gamma.png", plot = contour_b2_gamma_1, width = 10, height = 8)
ggsave(filename = "../Visualisation/HeteroUnif/contour_b1_b2.png", plot = contour_b1_b2_1, width = 10, height = 8)
ggsave(filename = "../Visualisation/HeteroUnif/contour_b1_d.png", plot = contour_b1_d_1, width = 10, height = 8)
ggsave(filename = "../Visualisation/HeteroUnif/contour_b2_d.png", plot = contour_b2_d_1, width = 10, height = 8)
ggsave(filename = "../Visualisation/HeteroUnif/contour_d_gamma.png", plot = contour_d_gamma_1, width = 10, height = 8)

```




```{r Plot the results 1}

## Plot the results

trace_plot_betas_1 = gg_trace_plot_betas(res_one, params, 10000, c(5000, 0.025, 7000, 0.0013), c(0, 0.03, 0, 0.0015))

trace_plot_gammad_1 = gg_trace_plot_gammad(res_one, params, 10000, c(3000, 0.025, 3000, 3), c(0.01, 0.25, 2.5, 7))

# aaa = gg_trace_plot_bp(res_one, params, 10000, c(20000, 0.025, 18000, 0.25), c(0, 0.03, 0, 0.3))

hist_plot_1 = gg_hist_plot(res_one, params, 10000, c(0.02, 90, 0.001, 1500, 0.2, 8, 8, 0.4), c(0, 0.03, 0, 0.0014, 0, 0.25, 2, 9))

hist_plot_w_prior_1 = gg_hist_plot_w_prior(res_one, params, 10000, c(1, 1, 1, 1, 1, 1, 0, 20), c(0.02, 90, 0.001, 1500, 0.2, 8, 8, 0.4), c(0, 0.03, 0, 0.0014, 0, 0.25, 2, 9))


## Save the results

ggsave(filename = "../Visualisation/HeteroUnif/trace_betas.png", plot = trace_plot_betas_1, width = 10, height = 8)

ggsave(filename = "../Visualisation/HeteroUnif/trace_gammad.png", plot = trace_plot_gammad_1, width = 10, height = 8)

ggsave(filename = "../Visualisation/HeteroUnif/hist_plot.png", plot = hist_plot_1, width = 10, height = 8)

ggsave(filename = "../Visualisation/HeteroUnif/hist_plot_w_prior.png", plot = hist_plot_w_prior_1, width = 10, height = 8)


```

```{r}

png("../Visualisation/HeteroUnif/autocorr.png", width=1000, height=800)
plotAuto(res_one[, c(2,3,4,5)], lag.max=50)
dev.off()

```


The MCMC algorithm for all parameters took `r round(end.time - start.time, 2)` minutes to run.

The algorithm estimates the mean, median, and quartiles of $\beta_1$ to be `r mean(res_one[-(1:5000),2])` (`r quantile(res_one[-(1:5000),2], probs = c(.25, .5, .75))`) compared to the true value of $\beta_1 =$  `r params[1]`, with an effective sample size of `r round(coda::effectiveSize(res_one[,2]))`. 

The algorithm estimates the mean, median, and quartiles of $\beta_2$ to be `r mean(res_one[-(1:5000),3])` (`r quantile(res_one[-(1:5000),3], probs = c(.25, .5, .75))`) compared to the true value of $\beta =$  `r params[2]`, with an effective sample size of `r round(coda::effectiveSize(res_one[,3]))`. 

It estimates the mean, median, and quartiles of $\gamma$ to be `r mean(res_one[-(1:5000),3])` (`r quantile(res_one[-(1:5000),3], probs = c(.25, .5, .75))`) compared to the true values of $\gamma =$ `r params[3]`, with an effective sample size of `r round(coda::effectiveSize(res_one[,3]))`.

It estimates the mean, median, and quartiles of $d$ to be `r mean(res_one[-(1:5000),4])` (`r quantile(res_one[-(1:5000),4], probs = c(.25, .5, .75))`) compared to the true values of $d =$ `r params[4]`, with an effective sample size of `r round(coda::effectiveSize(res_one[,4]))`.



# Making inference (Reparameterisation)

We now rerun the analysis under a reparameterised model in order to see if this has any effect on our ability to make efficient inference.

```{r Performing analysis reparam}

set.seed(1106)

## Run the algorithm

start.time <- Sys.time()

res.list.reparam_one <- MCMC_reparam(N.its = 100000, N = 100, inf.ids, rem.times, distance_mat, 
                     lambda.b1 = 1, nu.b1 = 1, lambda.g = 1 , nu.g = 1,
                     inc.beta1 = list(0.007, T), inc.p = list(0.00007, T), inc.dist = list(5, T),
                     inc.inf.times = list(NA, T), inc.gamma = list(0.11, T),
                     d.upper = 15, sigmap = 0.05, sigmad = 0.3, infupdate = 2)

res.reparam_one <- res.list.reparam_one[[1]]

end.time <- Sys.time()

print(c("Total Run Time:", end.time - start.time))

```

```{r}
coda::effectiveSize(res.reparam_one)

apply(res.reparam_one, 2, function(column) {
  quantiles <- quantile(column, c(0.025, 0.975))
  rounded_quantiles <- signif(quantiles, digits = 3)  # Round to 3 significant figures
  return(rounded_quantiles)
})


apply(res.reparam_one[ , ], 2, function(column) {
  std_dev <- sd(column)  # Calculate standard deviation
  rounded_std_dev <- signif(std_dev, digits = 3)  # Round to 3 significant figures
  return(rounded_std_dev)
})


```


```{r Summarise the results reparam 1}

## Summarise the results

summary(res.reparam_one[-(1:5000), 2:5])

Avg_acc_probs(res.list.reparam_one, T)

```

```{r}

hex_b1_gamma_reparam = gg_hex_plot_beta1_gamma(res.reparam_one[-(1:5000),], params_true, c(0.08,0.25,  0.08,0.22,  0.08,0.04,  0.08,0.01), c(0, 0.091), c(0, 0.26))

hex_b2_gamma_reparam = gg_hex_plot_beta2_gamma(res.reparam_one[-(1:5000),], params_true, c(0.0019,0.25,  0.0019,0.22,  0.0019,0.04,  0.0019,0.01), c(0, 0.0022), c(0, 0.26))

hex_b1_b2_reparam = gg_hex_plot_beta1_beta2(res.reparam_one[-(1:5000),], params_true, c(0.036,0.0017,  0.075,0.0017,  0.035,0.0019,  0.075,0.0019), c(0, 0.091), c(0, 0.0019))

hex_b1_d_reparam = gg_hex_plot_beta1_d(res.reparam_one[-(1:5000),], params_true, c(0.036,7.5,  0.075,7.5,  0.035,6.5,  0.075,6.5), c(0, 0.091), c(1.35, 8))

hex_b2_d_reparam = gg_hex_plot_beta2_d(res.reparam_one[-(1:5000),], params_true, c(0.00165,8,  0.00165,7.25,  0.0008,8,  0.0011,7.25), c(0, 0.0019), c(1.35, 8))

hex_d_gamma_reparam = gg_hex_plot_d_gamma(res.reparam_one[-(1:5000),], params_true, c(1.7,0,  3.2,0,  5,0,  7,0), c(1.35, 8), c(0, 0.26))


ggsave(filename = "../Visualisation/HeteroUnif/hex_b1_gamma_reparam.png", plot = hex_b1_gamma_reparam, width = 10, height = 8)
ggsave(filename = "../Visualisation/HeteroUnif/hex_b2_gamma_reparam.png", plot = hex_b2_gamma_reparam, width = 10, height = 8)
ggsave(filename = "../Visualisation/HeteroUnif/hex_b1_b2_reparam.png", plot = hex_b1_b2_reparam, width = 10, height = 8)
ggsave(filename = "../Visualisation/HeteroUnif/hex_b1_d_reparam.png", plot = hex_b1_d_reparam, width = 10, height = 8)
ggsave(filename = "../Visualisation/HeteroUnif/hex_b2_d_reparam.png", plot = hex_b2_d_reparam, width = 10, height = 8)
ggsave(filename = "../Visualisation/HeteroUnif/hex_d_gamma_reparam.png", plot = hex_d_gamma_reparam, width = 10, height = 8)

```



```{r}

contour_b1_gamma_reparam = gg_contour_plot_b1_gamma(res.reparam_one, params, 10000, c(0, 0.04), c(0, 0.175))

contour_b2_gamma_reparam = gg_contour_plot_b2_gamma(res.reparam_one, params, 10000, c(0, 0.0022), c(0, 0.175))

contour_b1_b2_reparam = gg_contour_plot_b1_b2(res.reparam_one, params, 10000, c(0, 0.015), c(0, 0.0019))

contour_b1_d_reparam = gg_contour_plot_b1_d(res.reparam_one, params, 10000, c(0, 0.015), c(3, 6.25))

contour_b2_d_reparam = gg_contour_plot_b2_d(res.reparam_one, params, 10000, c(0, 0.0019), c(4, 6.25))

contour_d_gamma_reparam = gg_contour_plot_d_gamma(res.reparam_one, params, 10000, c(4, 6.25), c(0, 0.26))


ggsave(filename = "../Visualisation/HeteroUnif/contour_b1_gamma_reparam.png", plot = contour_b1_gamma_reparam, width = 10, height = 8)
ggsave(filename = "../Visualisation/HeteroUnif/contour_b2_gamma_reparam.png", plot = contour_b2_gamma_reparam, width = 10, height = 8)
ggsave(filename = "../Visualisation/HeteroUnif/contour_b1_b2_reparam.png", plot = contour_b1_b2_reparam, width = 10, height = 8)
ggsave(filename = "../Visualisation/HeteroUnif/contour_b1_d_reparam.png", plot = contour_b1_d_reparam, width = 10, height = 8)
ggsave(filename = "../Visualisation/HeteroUnif/contour_b2_d_reparam.png", plot = contour_b2_d_reparam, width = 10, height = 8)
ggsave(filename = "../Visualisation/HeteroUnif/contour_d_gamma_reparam.png", plot = contour_d_gamma_reparam, width = 10, height = 8)

```


```{r Plot the results reparam 1}

## Plot the results

trace_plot_betas_reparam_1 = gg_trace_plot_betas(res.reparam_one, params, 10000, c(23000, 0.025, 23000, 0.0013), c(0, 0.03, 0, 0.0015))

trace_plot_gammad_reparam_1 = gg_trace_plot_gammad(res.reparam_one, params, 10000, c(3000, 0.025, 20000, 3), c(0.01, 0.25, 2.5, 7))

trace_plot_bp_reparam_1 = gg_trace_plot_bp(res.reparam_one, params, 10000, c(23000, 0.025, 27000, 0.175), c(0, 0.03, 0, 0.2))


hist_plot_betas_reparam_1 = gg_hist_plot(res.reparam_one, params, 10000, c(0.0175, 90, 0.001, 1500, 0.2, 8, 8, 0.4), c(0, 0.0225, 0, 0.00125, 0, 0.25, 2, 10))

hist_plot_bp_reparam_1 = gg_hist_plot_reparam(res.reparam_one, params, 10000, c(0.0175, 100, 0.1, 10, 0.2, 8, 8, 0.4), c(0, 0.0225, 0, 0.15, 0, 0.25, 2, 10))


hist_plot_w_prior_reparam = gg_hist_plot_w_prior(res.reparam_one, params, 10000, c(1, 1, 1, 1, 1, 1, 0, 20), c(0.0175, 90, 0.001, 1500, 0.2, 8, 8, 0.4), c(0, 0.0225, 0, 0.00125, 0, 0.25, 2, 10))

hist_plot_bp_w_prior_reparam = gg_hist_plot_w_prior_reparam(res.reparam_one, params, 10000, c(1, 1, 0, 1, 1, 1, 0, 20), c(0.0175, 100, 0.1, 10, 0.2, 8, 8, 0.4), c(0, 0.0225, 0, 0.15, 0, 0.25, 2, 10))

## Save the results

ggsave(filename = "../Visualisation/HeteroUnif/trace_betas_reparam.png", plot = trace_plot_betas_reparam_1, width = 10, height = 8)

ggsave(filename = "../Visualisation/HeteroUnif/trace_gammad_reparam.png", plot = trace_plot_gammad_reparam_1, width = 10, height = 8)

ggsave(filename = "../Visualisation/HeteroUnif/trace_bp_reparam.png", plot = trace_plot_bp_reparam_1, width = 10, height = 8)

ggsave(filename = "../Visualisation/HeteroUnif/hist_plot_betas_reparam.png", plot = hist_plot_betas_reparam_1, width = 10, height = 8)

ggsave(filename = "../Visualisation/HeteroUnif/hist_plot_bp_reparam.png", plot = hist_plot_bp_reparam_1, width = 10, height = 8)

ggsave(filename = "../Visualisation/HeteroUnif/hist_plot_w_prior_reparam.png", plot = hist_plot_w_prior_reparam, width = 10, height = 8)

ggsave(filename = "../Visualisation/HeteroUnif/hist_plot_bp_w_prior_reparam.png", plot = hist_plot_bp_w_prior_reparam, width = 10, height = 8)

```

```{r}

png("../Visualisation/HeteroUnif/autocorr_reparam_2.png", width=1000, height=800)
plotAuto(res.reparam_one[, c(2,3,4,5)], lag.max=50)
dev.off()

```

The MCMC algorithm for all parameters took `r round(end.time - start.time, 2)` minutes to run.

The algorithm estimates the mean, median, and quartiles of $\beta_1$ to be `r mean(res.reparam_one[-(1:5000),2])` (`r quantile(res.reparam_one[-(1:5000),2], probs = c(.25, .5, .75))`) compared to the true value of $\beta_1 =$  `r params[1]`, with an effective sample size of `r round(coda::effectiveSize(res.reparam_one[,2]))`. 

The algorithm estimates the mean, median, and quartiles of $\beta_2$ to be `r mean(res.reparam_one[-(1:5000),3])` (`r quantile(res.reparam_one[-(1:5000),3], probs = c(.25, .5, .75))`) compared to the true value of $\beta =$  `r params[2]`, with an effective sample size of `r round(coda::effectiveSize(res.reparam_one[,3]))`. 

It estimates the mean, median, and quartiles of $\gamma$ to be `r mean(res.reparam_one[-(1:5000),3])` (`r quantile(res.reparam_one[-(1:5000),3], probs = c(.25, .5, .75))`) compared to the true values of $\gamma =$ `r params[3]`, with an effective sample size of `r round(coda::effectiveSize(res.reparam_one[,3]))`.

It estimates the mean, median, and quartiles of $d$ to be `r mean(res.reparam_one[-(1:5000),4])` (`r quantile(res.reparam_one[-(1:5000),4], probs = c(.25, .5, .75))`) compared to the true values of $d =$ `r params[4]`, with an effective sample size of `r round(coda::effectiveSize(res.reparam_one[,4]))`.



