---
output:
  pdf_document: default
  html_document: default
title: "Unpaired Homogeneous General Stochastic Epidemic Inference"
author: "Benjamen Simon"
date: "September 2022"
knit: (function(inputFile, encoding) { 
          rmarkdown::render(
            inputFile,
            encoding = encoding, 
            output_dir = "../Reports") 
            })
header-includes:
  - \usepackage{amsmath}
  - \usepackage{enumerate}
  - \usepackage{bm}
  - \usepackage[T1]{fontenc}
---

```{r setup, include=FALSE}

# Rmarkdown settings
knitr::opts_chunk$set(echo = FALSE, cache = TRUE, eval= TRUE, message = FALSE, cache.path = "Caches/", fig.path = "Figures/HomoPaired/")

# Default to non-scientific notation
options(scipen=999)

# Suppress summarise info
options(dplyr.summarise.inform = FALSE)

```

```{r Packages}

  ## Data manipulation
  library(tidyverse)

  ## Plotting
  library(ggplot2)
  library(ggcorrplot)
  source("VisualisationFunctions.R")

  ## MCMC Diagnostics
  library(coda)

  ## Epidemics
  library(EpidemicR) # For simulating and making inference on GSE epidemics

```

# Load in the epidemic

The epidemic we have simulated has a modal number of infected individuals as roughly 25% of the population

```{r Load epidemic}

epi_list <- readRDS(file="../Data/Homo/epi_list.RData")

# Distance matrix
  spatial <- epi_list[[1]]
  xy.coords <- spatial[[1]]
  distance_mat <- spatial[[2]]

# Associated infection rate matrix
  params <- epi_list[[2]]
  rate_mat <- epi_list[[3]]
   
# Simulated epidemic
  Homo_sim <- epi_list[[4]]

# Total number of infected
  n_I = sum(Homo_sim[,2] < Inf)
  
# Who is infected
  inf.ids <- which(Homo_sim[,2] < Inf)

# Infection times
  inf.times <- Homo_sim[,2]

# Removal times
  rem.times <- Homo_sim[,3]
  
# Plot the epidemic
  epi_plot = EpidemicR::Plot_epidemic(xy.coords, Homo_sim, init.inf = c(1))

```




# Making inference (Infection times known)

Assuming we know the true infection times, we can remove the Metropolis-Hastings step and reduce the algorithm down to just a Gibbs sampler.

```{r Performing analysis I known}

set.seed(1104)

## Run the algorithm

start.time <- Sys.time()

res.list.Iknown <- MCMC_homo(N.its = 100000, N = 100, inf.ids, rem.times, distance_mat,
            inc.beta = list(NA, T), inc.inf.times = list(inf.times, F), inc.gamma = list(NA, T), infupdate = 6)

end.time <- Sys.time()

res.Iknown <- res.list.Iknown[[1]]


## Summarise the results

summary(res.Iknown[-(1:5000),])

```

```{r Plot the results I known}

## Plot the results

trace_plot_I_known = gg_trace_plot_homo(res.Iknown, params, 10000, c(20000, 0.0005, 20000, 0.03), c(0, 0.004, 0, 0.4))

hist_plot_I_known = gg_hist_plot_homo(res.Iknown, params, 10000, c(0.004, 700, 0.3, 8), c(0, 0.005, 0, 0.35))


## Save the results

ggsave(filename = "../Visualisation/HomoPaired/trace_I_known.png", plot = trace_plot_I_known, width = 10, height = 8)

ggsave(filename = "../Visualisation/HomoPaired/hist_I_known.png", plot = hist_plot_I_known, width = 10, height = 8)

```

The MCMC algorithm for all parameters took `r round(end.time - start.time, 2)` minutes to run.

The algorithm estimates the mean, median, and quartiles of $\beta$ to be `r mean(res.Iknown[-(1:5000),2])` (`r quantile(res.Iknown[-(1:5000),2], probs = c(.25, .5, .75))`) compared to the true value of $\beta =$  `r params[1]`, with an effective sample size of `r round(coda::effectiveSize(res.Iknown[,2]))`. It estimates the mean, median, and quartiles of $\gamma$ to be `r mean(res.Iknown[-(1:5000),3])` (`r quantile(res.Iknown[-(1:5000),3], probs = c(.25, .5, .75))`) compared to the true values of $\gamma =$ `r params[3]`, with an effective sample size of `r round(coda::effectiveSize(res.Iknown[,3]))`.



# Making inference (Infection times unknown)

Using our simulated epidemic from earlier, we will now apply our MCMC algorithm in an effort to make inference and recover the parameters.

```{r Performing analysis}

set.seed(1102)

## Run the algorithm

start.time <- Sys.time()

res.list <- MCMC_homo(N.its = 100000, N = 100, inf.ids, rem.times, distance_mat,
            inc.beta = list(NA, T), inc.inf.times = list(NA, T), inc.gamma = list(NA, T), infupdate = 6)

end.time <- Sys.time()

res.list[[2]] #avg. acceptance prob. of I

res <- res.list[[1]]

## Summarise the results

summary(res[-(1:5000),])

```

```{r Plot the results}

## Plot the results

trace_plot_I_unknown = gg_trace_plot_homo(res, params, 10000, c(20000, 0.0004, 20000, 0.04), c(0, 0.0065, 0, 0.65))

hist_plot_I_unknown = gg_hist_plot_homo(res, params, 10000, c(0.005, 350, 0.5, 4), c(0, 0.007, 0, 0.6))


## Save the results

ggsave(filename = "../Visualisation/HomoPaired/trace_I_unknown.png", plot = trace_plot_I_unknown, width = 10, height = 8)

ggsave(filename = "../Visualisation/HomoPaired/hist_I_unknown.png", plot = hist_plot_I_unknown, width = 10, height = 8)

```

The MCMC algorithm for all parameters took `r round(end.time - start.time, 2)` minutes to run.

The algorithm estimates the mean, median, and quartiles of $\beta$ to be `r mean(res[-(1:5000),2])` (`r quantile(res[-(1:5000),2], probs = c(.25, .5, .75))`) compared to the true value of $\beta =$ `r params[1]`, with an effective sample size of `r round(coda::effectiveSize(res[,2]))`. It estimates the mean, median, and quartiles of $\gamma$ to be `r mean(res[-(1:5000),3])` (`r quantile(res[-(1:5000),3], probs = c(.25, .5, .75))`) compared to the true values of $\gamma =$ `r params[3]`, with an effective sample size of `r round(coda::effectiveSize(res[,3]))`.



