---
output:
  pdf_document: default
  html_document: default
title: "Unpaired Homogeneous General Stochastic Epidemic Inference"
author: "Benjamen Simon"
date: "September 2022"
knit: (function(inputFile, encoding) { 
          rmarkdown::render(
            inputFile,
            encoding = encoding, 
            output_dir = "../Reports") 
            })
header-includes:
  - \usepackage{amsmath}
  - \usepackage{enumerate}
  - \usepackage{bm}
  - \usepackage[T1]{fontenc}
---

```{r setup, include=FALSE}

# Rmarkdown settings
knitr::opts_chunk$set(echo = FALSE, cache = TRUE, eval= TRUE, message = FALSE, cache.path = "Caches/", fig.path = "Figures/HomoPaired/")

# Default to non-scientific notation
options(scipen=999)

# Suppress summarise info
options(dplyr.summarise.inform = FALSE)

```

```{r Packages}

  ## Data manipulation
  library(tidyverse)

  ## Plotting
  library(ggplot2)
  library(ggcorrplot)
  source("VisualisationFunctions.R")
  library(plotMCMC) # autocorrelation plots and other diagnostic plots

  ## MCMC Diagnostics
  library(coda)

  ## Epidemics
  library(EpidemicR) # For simulating and making inference on GSE epidemics

```

# Load in the epidemic

The epidemic we have simulated has a modal number of infected individuals as roughly 25% of the population

```{r Load epidemic}

epi_list <- readRDS(file="../Data/Homo/epi_list.RData")

# Distance matrix
  spatial <- epi_list[[1]]
  xy.coords <- spatial[[1]]
  distance_mat <- spatial[[2]]

# Associated infection rate matrix
  params <- epi_list[[2]]
  rate_mat <- epi_list[[3]]
   
# Simulated epidemic
  Homo_sim <- epi_list[[4]]

# Total number of infected
  n_I = sum(Homo_sim[,2] < Inf)
  
# Who is infected
  inf.ids <- which(Homo_sim[,2] < Inf)

# Infection times
  inf.times <- Homo_sim[,2]

# Removal times
  rem.times <- Homo_sim[,3]
  
# Plot the epidemic
  epi_plot = EpidemicR::Plot_epidemic(xy.coords, Homo_sim, init.inf = c(1))

```


```{r}
createEpidemicDataFrame <- function(infection_times, removal_times, size_t) {
  # Create a sequence of time steps
  time_steps <- seq(0, max(removal_times) + size_t, by = size_t)
  
  # Initialize counts for S, I, and R states
  counts_S <- vector("integer", length = length(time_steps))
  counts_I <- vector("integer", length = length(time_steps))
  counts_R <- vector("integer", length = length(time_steps))
  
  # Initialize the initial state
  counts_S[1] <- 99
  counts_I[1] <- 1
  counts_R[1] <- 0
  
  # Loop through each time step
  for (i in 2:length(time_steps)) {
    t <- time_steps[i]
    
    # Count the number of individuals in each state at time t
    counts_S[i] <- counts_S[1] - sum(infection_times <= t)
    counts_I[i] <- counts_I[1] + sum(infection_times <= t) - sum(removal_times <= t)
    counts_R[i] <- counts_R[1] + sum(removal_times <= t)
  }
  
  # Create the data frame
  epidemic_df <- data.frame(
    Time = time_steps,
    S = counts_S,
    I = counts_I,
    R = counts_R
  )
  
  return(epidemic_df)
}
```

```{r}
epi_df_1 = createEpidemicDataFrame(inf.times[inf.times != Inf & inf.times > 0], rem.times[rem.times != Inf], 0.01)

epi_df_2 = createEpidemicDataFrame(inf.times[inf.times != Inf & inf.times > 0], rem.times[rem.times != Inf], 0.5)

epi_df_3 = createEpidemicDataFrame(inf.times[inf.times != Inf & inf.times > 0], rem.times[rem.times != Inf], 1)

epi_df_4 = createEpidemicDataFrame(inf.times[inf.times != Inf & inf.times > 0], rem.times[rem.times != Inf], 3)
```




```{r}

# Convert data to tidy format
tidy_df_1 <- tidyr::gather(epi_df_1, key = "variable", value = "value", -Time)

variable_order <- c("S", "I", "R")
color_order <- c("darkgreen", "red", "blue")

tidy_df_1$variable <- factor(tidy_df_1$variable, levels = variable_order)

# Create the plot
curves_1 = ggplot(tidy_df_1, aes(x = Time, y = value, color = variable)) +
  geom_step() +
  scale_color_manual(values = color_order, labels = variable_order) +
  ylim(0, 100)+
  ylab("Population")+
  ggtitle("Continuous")


ggsave(filename = "../Visualisation/curves_cont.png", plot = curves_1, width = 10, height = 8)


```




```{r}

# Convert data to tidy format
tidy_df_2 <- tidyr::gather(epi_df_2, key = "variable", value = "value", -Time)

variable_order <- c("S", "I", "R")
color_order <- c("darkgreen", "red", "blue")

tidy_df_2$variable <- factor(tidy_df_2$variable, levels = variable_order)

# Create the plot
curves_2 = ggplot(tidy_df_2, aes(x = Time, y = value, color = variable)) +
  geom_step() +
  scale_color_manual(values = color_order, labels = variable_order) +
  ylim(0, 100)+
  ylab("Population")+
  ggtitle(expression("Discrete:" ~ Delta * t == 0.5))


ggsave(filename = "../Visualisation/curves_2.png", plot = curves_2, width = 10, height = 8)


```




```{r}

# Convert data to tidy format
tidy_df_3 <- tidyr::gather(epi_df_3, key = "variable", value = "value", -Time)

variable_order <- c("S", "I", "R")
color_order <- c("darkgreen", "red", "blue")

tidy_df_3$variable <- factor(tidy_df_3$variable, levels = variable_order)

# Create the plot
curves_3 = ggplot(tidy_df_3, aes(x = Time, y = value, color = variable)) +
  geom_step() +
  scale_color_manual(values = color_order, labels = variable_order) +
  ylim(0, 100)+
  ylab("Population")+
  ggtitle(expression("Discrete:" ~ Delta * t == 1))


ggsave(filename = "../Visualisation/curves_3.png", plot = curves_3, width = 10, height = 8)


```




```{r}

# Convert data to tidy format
tidy_df_4 <- tidyr::gather(epi_df_4, key = "variable", value = "value", -Time)

variable_order <- c("S", "I", "R")
color_order <- c("darkgreen", "red", "blue")

tidy_df_4$variable <- factor(tidy_df_4$variable, levels = variable_order)

# Create the plot
curves_4 = ggplot(tidy_df_4, aes(x = Time, y = value, color = variable)) +
  geom_step() +
  scale_color_manual(values = color_order, labels = variable_order) +
  ylim(0, 100)+
  ylab("Population")+
  ggtitle(expression("Discrete:" ~ Delta * t == 3))


ggsave(filename = "../Visualisation/curves_4.png", plot = curves_4, width = 10, height = 8)


```





# Making inference (Infection times known)

Assuming we know the true infection times, we can remove the Metropolis-Hastings step and reduce the algorithm down to just a Gibbs sampler.

```{r Performing analysis I known}

set.seed(1104)

## Run the algorithm

start.time <- Sys.time()

res.list.Iknown <- MCMC_homo(N.its = 100000, N = 100, inf.ids, rem.times, distance_mat,
                             lambda.b = 1, nu.b = 1, lambda.g = 1 , nu.g = 1,
            inc.beta = list(NA, T), inc.inf.times = list(inf.times, F), inc.gamma = list(NA, T), infupdate = 6)

end.time <- Sys.time()

res.Iknown <- res.list.Iknown[[1]]


## Summarise the results

summary(res.Iknown[-(1:5000),])

```

```{r}

hex_beta_delta_I_known = gg_hex_plot_homo(res.Iknown[-c(1:10000), ], params_true, c(0.007,0.1,  0.007,0.05,  0.007,0.53,  0.007,0.6), c(0, 0.008), c(0, 0.65))

hex_beta_delta_I_known

ggsave(filename = "../Visualisation/HomoPaired/hex_beta_delta_I_known.png", plot = hex_beta_delta_I_known, width = 10, height = 8)

```




```{r Plot the results I known}

## Plot the results

trace_plot_I_known = gg_trace_plot_homo(res.Iknown, params, 10000, c(20000, 0.0005, 20000, 0.03), c(0, 0.004, 0, 0.4))

hist_plot_I_known = gg_hist_plot_homo(res.Iknown, params, 10000, c(0.004, 700, 0.3, 8), c(0, 0.005, 0, 0.35))

hist_plot_w_prior_I_known = gg_hist_plot_w_prior_homo(res.Iknown, params, 10000, c(1, 1, 1, 1), c(0.004, 700, 0.3, 8), c(0, 0.005, 0, 0.35))


## Save the results

ggsave(filename = "../Visualisation/HomoPaired/trace_I_known.png", plot = trace_plot_I_known, width = 10, height = 8)

ggsave(filename = "../Visualisation/HomoPaired/hist_I_known.png", plot = hist_plot_I_known, width = 10, height = 8)

ggsave(filename = "../Visualisation/HomoPaired/hist_w_prior_I_known.png", plot = hist_plot_w_prior_I_known, width = 10, height = 8)

```

The MCMC algorithm for all parameters took `r round(end.time - start.time, 2)` minutes to run.

The algorithm estimates the mean, median, and quartiles of $\beta$ to be `r mean(res.Iknown[-(1:5000),2])` (`r quantile(res.Iknown[-(1:5000),2], probs = c(.25, .5, .75))`) compared to the true value of $\beta =$  `r params[1]`, with an effective sample size of `r round(coda::effectiveSize(res.Iknown[,2]))`. It estimates the mean, median, and quartiles of $\delta$ to be `r mean(res.Iknown[-(1:5000),3])` (`r quantile(res.Iknown[-(1:5000),3], probs = c(.25, .5, .75))`) compared to the true values of $\delta =$ `r params[3]`, with an effective sample size of `r round(coda::effectiveSize(res.Iknown[,3]))`.


```{r}

png("../Visualisation/HomoPaired/autocorr_Iknown.png", width=1000, height=800)
plotAuto(res.Iknown[, c(2,3)], lag.max=50)
dev.off()

```









# Making inference (Infection times unknown)

Using our simulated epidemic from earlier, we will now apply our MCMC algorithm in an effort to make inference and recover the parameters.

```{r Performing analysis}

set.seed(1102)

## Run the algorithm

start.time <- Sys.time()

res.list <- MCMC_homo(N.its = 100000, N = 100, inf.ids, rem.times, distance_mat,
                      lambda.b = 1, nu.b = 1, lambda.g = 1 , nu.g = 1,
                      inc.beta = list(NA, T), inc.inf.times = list(NA, T), inc.gamma = list(NA, T), infupdate = 6)

end.time <- Sys.time()

res <- res.list[[1]]

## Summarise the results

print(c("acceptance probability = ", res.list[[2]]))

summary(res[-(1:5000),])

```


```{r}

hex_beta_delta_I_unknown = gg_hex_plot_homo(res[-c(1:10000), ], params_true, c(0.007,0.1,  0.007,0.05,  0.001,0.53,  0.001,0.6), c(0, 0.008), c(0, 0.65))

hex_beta_delta_I_unknown

ggsave(filename = "../Visualisation/HomoPaired/hex_beta_delta_I_unknown.png", plot = hex_beta_delta_I_unknown, width = 10, height = 8)

```

```{r}

contour_beta_delta_I_unknown = gg_contour_plot_homo(res, params, 10000,  c(0, 0.008), c(0, 0.4))

contour_beta_delta_I_unknown

ggsave(filename = "../Visualisation/HomoPaired/contour_beta_delta_I_unknown.png", plot = contour_beta_delta_I_unknown, width = 10, height = 8)

```




```{r Plot the results}

## Plot the results

trace_plot_I_unknown = gg_trace_plot_homo(res, params, 10000, c(20000, 0.0004, 20000, 0.04), c(0, 0.0065, 0, 0.65))

hist_plot_I_unknown = gg_hist_plot_homo(res, params, 10000, c(0.005, 350, 0.5, 4), c(0, 0.007, 0, 0.6))

hist_plot_w_prior_I_unknown = gg_hist_plot_w_prior_homo(res, params, 10000, c(1, 1, 1, 1), c(0.006, 350, 0.6, 4), c(0, 0.007, 0, 0.7))


## Save the results

ggsave(filename = "../Visualisation/HomoPaired/trace_I_unknown.png", plot = trace_plot_I_unknown, width = 10, height = 8)

ggsave(filename = "../Visualisation/HomoPaired/hist_I_unknown.png", plot = hist_plot_I_unknown, width = 10, height = 8)

ggsave(filename = "../Visualisation/HomoPaired/hist_w_prior_I_unknown.png", plot = hist_plot_w_prior_I_unknown, width = 10, height = 8)

```

The MCMC algorithm for all parameters took `r round(end.time - start.time, 2)` minutes to run.

The algorithm estimates the mean, median, and quartiles of $\beta$ to be `r mean(res[-(1:5000),2])` (`r quantile(res[-(1:5000),2], probs = c(.25, .5, .75))`) compared to the true value of $\beta =$ `r params[1]`, with an effective sample size of `r round(coda::effectiveSize(res[,2]))`. It estimates the mean, median, and quartiles of $\delta$ to be `r mean(res[-(1:5000),3])` (`r quantile(res[-(1:5000),3], probs = c(.25, .5, .75))`) compared to the true values of $\delta =$ `r params[3]`, with an effective sample size of `r round(coda::effectiveSize(res[,3]))`.


```{r}

png("../Visualisation/HomoPaired/autocorr_Iunknown.png", width=1000, height=800)
plotAuto(res[, c(2,3)], lag.max=50)
dev.off()

```
