---
output:
  pdf_document: default
  html_document: default
title: "Generate Simulated Epidemic Data"
author: "Benjamen Simon"
date: "September 2022"
knit: (function(inputFile, encoding) { 
          rmarkdown::render(
            inputFile,
            encoding = encoding, 
            output_dir = "../Reports") 
            })
header-includes:
  - \usepackage{amsmath}
  - \usepackage{enumerate}
  - \usepackage{bm}
  - \usepackage[T1]{fontenc}
---

```{r setup, include=FALSE}

# Rmarkdown settings
knitr::opts_chunk$set(echo = FALSE, cache = TRUE, eval= TRUE, message = FALSE, cache.path = "Caches/", fig.path = "Figures/")

# Default to non-scientific notation
options(scipen=999)

# Suppress summarise info
options(dplyr.summarise.inform = FALSE)

```

```{r Packages}

  ## Data manipulation
  library(tidyverse)

  ## Epidemics
  library(EpidemicR) # For simulating and making inference on GSE epidemics

```

# Custom Functions

## Wrapper Function for generating a single epidemic

```{r Wrapper function}

GenerateEpidemic <- function(model, N, betas, d, gamma, ...){
  
  if(model == "Unif"){
    distfunc <- EpidemicR::Dist_mat_unif
  }
  if(model == "Clust"){
    distfunc <- EpidemicR::Dist_mat_clust
  }
  if(model == "Edge"){
    distfunc <- EpidemicR::Dist_mat_edge
  }
  if(model == "Latt"){
    distfunc <- EpidemicR::Dist_mat_latt
  }
  if(model == "Lines"){
    distfunc <- EpidemicR::Dist_mat_lines
  }
  if(model == "Trials"){
    distfunc <- EpidemicR::Dist_mat_trials
  }
  
  # Generate a distance matrix
   distance_mat <- distfunc(N, ...)[[2]]

  # Generate an associated infection rate matrix
   rate_mat <- EpidemicR::Beta_mat_form(distance_mat, betas, d) # A slight trick is used to generate a homogeneous epidemic
   
  # Generate a simulated epidemic
   sim <- EpidemicR::GSE_sim(N = N, beta.mat = rate_mat, gamma = gamma)

  return(sim)
}

# GenerateEpidemic(model = "Unif", N = 100, betas = c(0.002, 0.002), d = 10, gamma = 0.15, xlim = 20, ylim = 20)

```

## Function for idenitifying a parameter set

```{r Parameter Set Identifier}

Identify_parameter_set <- function(model, N, betas, ps, gammas, ds, nsims, ...){
  
  num_param_sets = (length(betas)*length(ps)*length(gammas)*length(ds))

  n_I = array(NA, c(nsims*num_param_sets, 5))
  # one row for each sim, one column per variable of interest
  
  counter = 1
  
  for(b in betas){
    for(p in ps){
      for(d in ds){
        for(g in gammas){
          for(s in 1:nsims){
            
            sim = GenerateEpidemic(model = model, N = N, betas = c(b, p*b), d = d, gamma = g, ...)
            
            n_I[counter, c(1,2,3,4)] = c(b, p*b, d, g)
            n_I[counter, 5] = sum(sim[,2] < Inf)/N * 100
            
            counter = counter + 1
          }
        }
      }
    }
  }
  
  nI_df = data.frame(n_I) %>% 
                    rename(beta_1 = X1, beta_2 = X2, d = X3, gamma = X4, nI = X5) %>% 
                    group_by(beta_1, beta_2, d, gamma) %>% 
                    summarise(mean = mean(nI), median = median(nI), in_range = sum(nI > 20 & nI < 30)) %>% 
                    filter(quantile(in_range, 0.9)<in_range) %>% 
                    arrange(desc(in_range), mean, median)
  
  return(nI_df)
}

# Identify_parameter_set(model = "Unif", N = 100, betas = seq(0.001, 0.006, 0.001), ps = c(1), ds = c(10), gammas = c(0.12), nsims = 10, xlim = 20, ylim = 20)

```

## Function to generate the histogram of final size

```{r Histogram of final size}

Hist_final_size <- function(model, N, betas, gamma, d, nsims, ...){
  
  n_I = rep(NA, nsims)
  
  valid_seeds = array(NA, nsims)
  
  for(s in 1:nsims){
    
    set.seed(s)
    
    sim = GenerateEpidemic(model = model, N = N, betas = betas, d = d, gamma = gamma, ...)
    
    n_I[s] = sum(sim[,2] < Inf)/N * 100
    
    if(n_I[s] == 25){
      
      valid_seeds[s] = s
    }
  }
  
  print(head(na.omit(as.data.frame(valid_seeds))))

  hist(n_I, breaks = 50)
  abline(v = mean(n_I), # Add line for mean
          col = "red",
          lwd = 3)
  text(x = mean(n_I)+ 11, # Add text for mean
     y = mean(n_I) * nsims * 0.01,
     paste("Mean =", mean(n_I)),
     col = "red",
     cex = 1)
  abline(v = median(n_I),# Add line for median
         col = "blue",
         lwd = 3)
  text(x = median(n_I)+ 9, # Add text for median
     y = median(n_I) * nsims * 0.1,
     paste("Median =", median(n_I)),
     col = "blue",
     cex = 1)
}

# Hist_final_size(model = "Unif", N = 100, betas = c(0.0005, 0.0005), d = 10, gamma = 0.03, nsims = 500, xlim = 20, ylim = 20)

```

# Homogeneous Epidemic

The case where all individuals in the population make contact at a single rate $\beta$.

First identify a parameter set with an modal total number infected of around 25% of the population.

```{r Identify a parameter set for Homogeneous epidemic}

set.seed(1)

betas = seq(0.001, 0.005, 0.0005)

gammas = seq(0.1, 0.3, 0.01)

parameter_options = Identify_parameter_set(model = "Unif", N = 100, betas = betas, ps = c(1), ds = c(10), gammas = gammas, nsims = 1000, xlim = 20, ylim = 20)

parameter_options

```

In this case we choose $\beta = 0.003$ and $\gamma = 0.23$.

The following histograms show the distribution of final sizes for this parameter set, and the table provides some seeds that generate an epidemic of the desired size.

```{r Plot histograms for Homogeneous epidemic}

Hist_final_size(model = "Unif", N = 100, betas = c(0.0023, 0.0023), d = 10, gamma = 0.18, nsims = 10000, xlim = 20, ylim = 20)

# Hist_final_size(model = "Unif", N = 100, betas = c(0.003, 0.003), d = 10, gamma = 0.23, nsims = 10000, xlim = 20, ylim = 20)
# 
# Hist_final_size(model = "Unif", N = 100, betas = c(0.005, 0.005), d = 10, gamma = 0.29, nsims = 10000, xlim = 20, ylim = 20)

```

Finally, generate the chosen epidemic and save the results

```{r Generate a Homogeneous epidemic}

set.seed(263)

# Generate a distance matrix
  spatial <- EpidemicR::Dist_mat_unif(N=100, xlim = 20, ylim = 20)
  xy.coords <- spatial[[1]]
  distance_mat <- spatial[[2]]

# Generate an associated infection rate matrix
  params <- c("beta1" = 0.003, "beta2" = 0.003, "gamma" = 0.23, "d" = 10, "p" = 1, "llh" = NA)
  rate_mat <- EpidemicR::Beta_mat_form(distance_mat, c(params["beta1"], params["beta2"]), params["d"]) # A slight trick is used to generate a homogeneous epidemic
   
# Generate a simulated epidemic
  Homo_sim <- EpidemicR::GSE_sim(N = 100, beta.mat = rate_mat, gamma = 0.23)

# Total number of infected
  n_I = sum(Homo_sim[,2] < Inf)
  
# Who is infected
  inf.ids <- which(Homo_sim[,2] < Inf)

# Infection times
  inf.times <- Homo_sim[,2]

# Removal times
  rem.times <- Homo_sim[,3]
  
# Calculate Likelihood
  params["llh"] <- EpidemicR::log_likelihood(inf.times, rem.times, rate_mat)
  
# Plot the epidemic
  pdf("../Data/Homo/epi_plot.pdf") 
  epi_plot = EpidemicR::Plot_epidemic(xy.coords, Homo_sim, init.inf = c(1))
  while (!is.null(dev.list()))  dev.off() # save plot
```

```{r Save the Homogeneous epidemic}

epi_list = list(spatial, params, rate_mat, Homo_sim)

saveRDS(epi_list, file="../Data/Homo/epi_list.RData")

```





# Heterogeneous Epidemic (Uniform spatial distribution)

The case where individuals are arranged randomly on a grid, and individuals within Euclidean distance $d$ of each other make contact at rate $\beta_1$, otherwise at rate $\beta_2$.

First identify a parameter set with an modal total number infected of around 25% of the population.

```{r Identify a parameter set for Heterogeneous (Unif) epidemic}

set.seed(1)

betas = seq(0.001, 0.007, 0.0005)

ps = seq(0.01, 0.1, 0.01)

gammas = seq(0.1, 0.4, 0.01)

ds = seq(1, 5, 1)

parameter_options = Identify_parameter_set(model = "Unif", N = 100, betas = betas, ps = ps, ds = ds, gammas = gammas, nsims = 1000, xlim = 20, ylim = 20)

parameter_options

```

In this case we choose $\beta_1 = 0.007$, $\beta_2 = 0.00007$, $d = 5$, and $\gamma = 0.11$.

The following histograms show the distribution of final sizes for this parameter set, and the table provides some seeds that generate an epidemic of the desired size.

```{r Plot histograms for Heterogeneous (Unif) epidemic}

Hist_final_size(model = "Unif", N = 100, betas = c(0.007, 0.00007), d = 5, gamma = 0.11, nsims = 10000, xlim = 20, ylim = 20)

```

Finally, generate the chosen epidemic and save the results

```{r Generate epidemic 2}

set.seed(295)

# Generate a distance matrix
   spatial <- EpidemicR::Dist_mat_unif(N=100, xlim = 20, ylim = 20)
   xy.coords <- spatial[[1]]
   distance_mat <- spatial[[2]]

# Generate an associated infection rate matrix
   params <- c("beta1" = 0.007, "beta2" = 0.00007, "gamma" = 0.11, "d" = 5, "p" = 0.01, "llh" = NA)
   rate_mat <- EpidemicR::Beta_mat_form(distance_mat, c(params["beta1"], params["beta2"]), params["d"])
   
# Generate a simulated epidemic
   Hetero_sim <- EpidemicR::GSE_sim(N = 100, beta.mat = rate_mat, gamma = 0.11)
   
# Total number of infected
   n_I = sum(Hetero_sim[,2] < Inf)
  
# Who is infected
   inf.ids <- which(Hetero_sim[,2] < Inf)

# Infection times
  inf.times <- Hetero_sim[,2]

# Removal times
  rem.times <- Hetero_sim[,3]

# Calculate Likelihood
  params["llh"] <- EpidemicR::log_likelihood(inf.times, rem.times, rate_mat)
  
# Plot the epidemic
  pdf("../Data/HeteroUnif/epi_plot.pdf") 
  epi_plot = EpidemicR::Plot_epidemic(xy.coords, Hetero_sim, init.inf = c(1))
  while (!is.null(dev.list()))  dev.off() # save plot
```

```{r Save the Hetero Unif epidemic}

epi_list = list(spatial, params, rate_mat, Hetero_sim)

saveRDS(epi_list, file="../Data/HeteroUnif/epi_list.RData")

```




# Heterogeneous Epidemic (Uniform spatial distribution with $\beta_2$ = 0)

The case where individuals are arranged randomly on a grid, and individuals within Euclidean distance $d$ of each other make contact at rate $\beta_1$, otherwise at rate $\beta_2 = 0$, ie. they don't make contact.

First identify a parameter set with an modal total number infected of around 25% of the population.

```{r Identify a parameter set for Heterogeneous (b2 = 0) epidemic}

set.seed(1)

betas = seq(0.001, 0.007, 0.0005)

gammas = seq(0.1, 0.4, 0.02)

ds = seq(0.5, 5, 0.5)

parameter_options = Identify_parameter_set(model = "Unif", N = 100, betas = betas, ps = c(0), ds = ds, gammas = gammas, nsims = 1000, xlim = 20, ylim = 20)

parameter_options

```

In this case we choose $\beta_1 = 0.007$, $\beta_2 = 0$, $d = 5$, and $\gamma = 0.1$.

The following histograms show the distribution of final sizes for this parameter set, and the table provides some seeds that generate an epidemic of the desired size.

```{r Plot histograms for Heterogeneous (b2 = 0) epidemic}

Hist_final_size(model = "Unif", N = 100, betas = c(0.007, 0), d = 10, gamma = 0.1, nsims = 10000, xlim = 50, ylim = 10)

```

Finally, generate the chosen epidemic and save the results

```{r Generate epidemic 3}

set.seed(13)

# Generate a distance matrix
   spatial <- EpidemicR::Dist_mat_unif(N=100, xlim = 50, ylim = 10)
   xy.coords <- spatial[[1]]
   distance_mat <- spatial[[2]]

# Generate an associated infection rate matrix
   params <- c("beta1" = 0.1, "beta2" = 0, "gamma" = 0.15, "d" = 2.5, "p" = 0, "llh" = NA)
   rate_mat <- EpidemicR::Beta_mat_form(distance_mat, c(params["beta1"], params["beta2"]), params["d"])
   
# Generate a simulated epidemic
   Hetero_sim <- EpidemicR::GSE_sim(N = 100, beta.mat = rate_mat, gamma = 0.15)
   
# Total number of infected
   n_I = sum(Hetero_sim[,2] < Inf)
  
# Who is infected
   inf.ids <- which(Hetero_sim[,2] < Inf)

# Infection times
  inf.times <- Hetero_sim[,2]

# Removal times
  rem.times <- Hetero_sim[,3]

# Calculate Likelihood
  params["llh"] <- EpidemicR::log_likelihood(inf.times, rem.times, rate_mat)
  
# Plot the epidemic
  pdf("../Data/Hetero_b2_0/epi_plot.pdf") 
  epi_plot = EpidemicR::Plot_epidemic(xy.coords, Hetero_sim, init.inf = c(1))
  while (!is.null(dev.list()))  dev.off() # save plot
```

```{r Save the Hetero b2 0 epidemic}

epi_list = list(spatial, params, rate_mat, Hetero_sim)

saveRDS(epi_list, file="../Data/Hetero_b2_0/epi_list.RData")

```





# Heterogeneous Epidemic (Uniform spatial distribution on a rectangular space)

The case where individuals are arranged randomly on a grid, and individuals within Euclidean distance $d$ of each other make contact at rate $\beta_1$, otherwise at rate $\beta_2$. This time the grid is a long thin rectangle instead of a square.

First identify a parameter set with an modal total number infected of around 25% of the population.

```{r Identify a parameter set for Heterogeneous (Rect) epidemic}

set.seed(1)

betas = seq(0.001, 0.007, 0.0005)

ps = seq(0.01, 0.1, 0.01)

gammas = seq(0.1, 0.4, 0.02)

ds = seq(0.5, 5, 0.5)

parameter_options = Identify_parameter_set(model = "Unif", N = 100, betas = betas, ps = ps, ds = ds, gammas = gammas, nsims = 1000, xlim = 50, ylim = 10)

parameter_options

```

In this case we choose $\beta_1 = 0.0020$, $\beta_2 = 0.00007$, $d = 5$, and $\gamma = 0.11$.

The following histograms show the distribution of final sizes for this parameter set, and the table provides some seeds that generate an epidemic of the desired size.

```{r Plot histograms for Heterogeneous (Rect) epidemic}

Hist_final_size(model = "Unif", N = 100, betas = c(0.003, 0.003), d = 10, gamma = 0.23, nsims = 10000, xlim = 50, ylim = 10)

Hist_final_size(model = "Unif", N = 100, betas = c(0.005, 0.005), d = 10, gamma = 0.29, nsims = 10000, xlim = 50, ylim = 10)

```

Finally, generate the chosen epidemic and save the results

```{r Generate epidemic 4}

set.seed(13)

# Generate a distance matrix
   spatial <- EpidemicR::Dist_mat_unif(N=100, xlim = 50, ylim = 10)
   xy.coords <- spatial[[1]]
   distance_mat <- spatial[[2]]

# Generate an associated infection rate matrix
   params <- c("beta1" = 0.1, "beta2" = 0, "gamma" = 0.15, "d" = 2.5, "p" = 0, "llh" = NA)
   rate_mat <- EpidemicR::Beta_mat_form(distance_mat, c(params["beta1"], params["beta2"]), params["d"])
   
# Generate a simulated epidemic
   Hetero_sim <- EpidemicR::GSE_sim(N = 100, beta.mat = rate_mat, gamma = 0.15)
   
# Total number of infected
   n_I = sum(Hetero_sim[,2] < Inf)
  
# Who is infected
   inf.ids <- which(Hetero_sim[,2] < Inf)

# Infection times
  inf.times <- Hetero_sim[,2]

# Removal times
  rem.times <- Hetero_sim[,3]

# Calculate Likelihood
  params["llh"] <- EpidemicR::log_likelihood(inf.times, rem.times, rate_mat)
  
# Plot the epidemic
  pdf("../Data/HeteroRect/epi_plot.pdf") 
  epi_plot = EpidemicR::Plot_epidemic(xy.coords, Hetero_sim, init.inf = c(1))
  while (!is.null(dev.list()))  dev.off() # save plot
```

```{r Save the Hetero Rect epidemic}

epi_list = list(spatial, params, rate_mat, Hetero_sim)

saveRDS(epi_list, file="../Data/HeteroRect/epi_list.RData")

```







# Heterogeneous Epidemic in a Big Population (Uniform spatial distribution)

The case where individuals are arranged randomly on a grid, and individuals within Euclidean distance $d$ of each other make contact at rate $\beta_1$, otherwise at rate $\beta_2$. In this case the population is 500 times larger.

First identify a parameter set with an modal total number infected of around 25% of the population.

```{r Identify a parameter set for Heterogeneous (Big) epidemic}

set.seed(1)

betas = seq(0.003, 0.007, 0.0005)

ps = seq(0.01, 0.05, 0.01)

gammas = seq(0.2, 0.4, 0.02)

ds = seq(1, 5, 1)

parameter_options = Identify_parameter_set(model = "Unif", N = 50000, betas = betas, ps = ps, ds = ds, gammas = gammas, nsims = 1000, xlim = 20, ylim = 20)

parameter_options

```

In this case we choose $\beta_1 = 0.007$, $\beta_2 = 0.00007$, $d = 5$, and $\gamma = 0.11$.

The following histograms show the distribution of final sizes for this parameter set, and the table provides some seeds that generate an epidemic of the desired size.

```{r Plot histograms for Heterogeneous (Big) epidemic}

Hist_final_size(model = "Unif", N = 500, betas = c(0.007, 0.00007), d = 5, gamma = 0.11, nsims = 10000, xlim = 20, ylim = 20)

```

Finally, generate the chosen epidemic and save the results

```{r Generate epidemic 5}

set.seed(295)

# Generate a distance matrix
   spatial <- EpidemicR::Dist_mat_unif(N=50000, xlim = 20, ylim = 20)
   xy.coords <- spatial[[1]]
   distance_mat <- spatial[[2]]

# Generate an associated infection rate matrix
   params <- c("beta1" = 0.007, "beta2" = 0.00007, "gamma" = 0.11, "d" = 5, "p" = 0.01, "llh" = NA)
   rate_mat <- EpidemicR::Beta_mat_form(distance_mat, c(params["beta1"], params["beta2"]), params["d"])
   
# Generate a simulated epidemic
   Hetero_sim <- EpidemicR::GSE_sim(N = 100, beta.mat = rate_mat, gamma = 0.11)
   
# Total number of infected
   n_I = sum(Hetero_sim[,2] < Inf)
  
# Who is infected
   inf.ids <- which(Hetero_sim[,2] < Inf)

# Infection times
  inf.times <- Hetero_sim[,2]

# Removal times
  rem.times <- Hetero_sim[,3]

# Calculate Likelihood
  params["llh"] <- EpidemicR::log_likelihood(inf.times, rem.times, rate_mat)
  
# Plot the epidemic
  pdf("../Data/HeteroBig/epi_plot.pdf") 
  epi_plot = EpidemicR::Plot_epidemic(xy.coords, Hetero_sim, init.inf = c(1))
  while (!is.null(dev.list()))  dev.off() # save plot
```

```{r Save the Hetero Big epidemic}

epi_list = list(spatial, params, rate_mat, Hetero_sim)

saveRDS(epi_list, file="../Data/HeteroBig/epi_list.RData")

```






# Heterogeneous Epidemic (Clustered spatial distribution)


The case where individuals are distributed using a cluster process. The poles of each cluster were distributed uniformly on a $20\times20$ plane, and the rest of the population were distributed evenly around each pole using an isotropic Gaussian distribution. Individuals within Euclidean distance $d$ of each other make contact at rate $\beta_1$, otherwise at rate $\beta_2$.

First identify a parameter set with an modal total number infected of around 25% of the population.

```{r Identify a parameter set for Heterogeneous (Clust) epidemic}

set.seed(1)

betas = seq(0.001, 0.007, 0.0005)

ps = seq(0.01, 0.1, 0.01)

gammas = seq(0.1, 0.4, 0.02)

ds = seq(0.5, 5, 0.5)

parameter_options = Identify_parameter_set(model = "Clust", N = 100, betas = betas, ps = ps, ds = ds, gammas = gammas, nsims = 1000, xlim = 20, ylim = 20, centres = 5, spread = 1)

parameter_options

```

In this case we choose $\beta_1 = 0.007$, $\beta_2 = 0.00007$, $d = 5$, and $\gamma = 0.11$.

The following histograms show the distribution of final sizes for this parameter set, and the table provides some seeds that generate an epidemic of the desired size.

```{r Plot histograms for Heterogeneous (Clust) epidemic}

Hist_final_size(model = "Clust", N = 100, betas = c(0.007, 0.00007), d = 5, gamma = 0.11, nsims = 10000, xlim = 20, ylim = 20, centres = 5, spread = 1)

```

Finally, generate the chosen epidemic and save the results

```{r Generate epidemic 6}

set.seed(295)

# Generate a distance matrix
   spatial <- EpidemicR::Dist_mat_clust(N=100, xlim = 20, ylim = 20, centres = 5, spread = 1)
   xy.coords <- spatial[[1]]
   distance_mat <- spatial[[2]]

# Generate an associated infection rate matrix
   params <- c("beta1" = 0.007, "beta2" = 0.00007, "gamma" = 0.11, "d" = 5, "p" = 0.01, "llh" = NA)
   rate_mat <- EpidemicR::Beta_mat_form(distance_mat, c(params["beta1"], params["beta2"]), params["d"])
   
# Generate a simulated epidemic
   Hetero_sim <- EpidemicR::GSE_sim(N = 100, beta.mat = rate_mat, gamma = 0.11)
   
# Total number of infected
   n_I = sum(Hetero_sim[,2] < Inf)
  
# Who is infected
   inf.ids <- which(Hetero_sim[,2] < Inf)

# Infection times
  inf.times <- Hetero_sim[,2]

# Removal times
  rem.times <- Hetero_sim[,3]

# Calculate Likelihood
  params["llh"] <- EpidemicR::log_likelihood(inf.times, rem.times, rate_mat)
  
# Plot the epidemic
  pdf("../Data/HeteroClust/epi_plot.pdf") 
  epi_plot = EpidemicR::Plot_epidemic(xy.coords, Hetero_sim, init.inf = c(1))
  while (!is.null(dev.list()))  dev.off() # save plot
```

```{r Save the Hetero Clust epidemic}

epi_list = list(spatial, params, rate_mat, Hetero_sim)

saveRDS(epi_list, file="../Data/HeteroClust/epi_list.RData")

```







# Heterogeneous Epidemic (Lattice spatial distribution)

The case where half the population are arranged on a lattice. Each individual then has a `buddy' which is placed around them using an isotropic Gaussian distribution, and individuals within Euclidean distance $d$ of each other make contact at rate $\beta_1$, otherwise at rate $\beta_2$.

First identify a parameter set with an modal total number infected of around 25% of the population.

```{r Identify a parameter set for Heterogeneous (Latt) epidemic}

set.seed(1)

betas = seq(0.001, 0.007, 0.0005)

ps = seq(0.01, 0.1, 0.01)

gammas = seq(0.1, 0.4, 0.02)

ds = seq(0.5, 5, 0.5)

parameter_options = Identify_parameter_set(model = "Latt", N = 100, betas = betas, ps = ps, ds = ds, gammas = gammas, nsims = 1000, xlim = 20, ylim = 20, buddy = TRUE)

parameter_options

```

In this case we choose $\beta_1 = 0.007$, $\beta_2 = 0.00007$, $d = 5$, and $\gamma = 0.11$.

The following histograms show the distribution of final sizes for this parameter set, and the table provides some seeds that generate an epidemic of the desired size.

```{r Plot histograms for Heterogeneous (Latt) epidemic}

Hist_final_size(model = "Latt", N = 200, betas = c(0.007, 0.00007), d = 5, gamma = 0.11, nsims = 10000, xlim = 20, ylim = 20, nrow = 10, buddy = TRUE)

```

Finally, generate the chosen epidemic and save the results

```{r Generate epidemic 7}

set.seed(295)

# Generate a distance matrix
   spatial <- Dist_mat_latt(N = 200, nrow = 10, xlim = 20, ylim = 20, buddy = TRUE)
   xy.coords <- spatial[[1]]
   distance_mat <- spatial[[2]]

# Generate an associated infection rate matrix
   params <- c("beta1" = 0.007, "beta2" = 0.00007, "gamma" = 0.11, "d" = 5, "p" = 0.01, "llh" = NA)
   rate_mat <- EpidemicR::Beta_mat_form(distance_mat, c(params["beta1"], params["beta2"]), params["d"])
   
# Generate a simulated epidemic
   Hetero_sim <- EpidemicR::GSE_sim(N = 200, beta.mat = rate_mat, gamma = 0.11)
   
# Total number of infected
   n_I = sum(Hetero_sim[,2] < Inf)
  
# Who is infected
   inf.ids <- which(Hetero_sim[,2] < Inf)

# Infection times
  inf.times <- Hetero_sim[,2]

# Removal times
  rem.times <- Hetero_sim[,3]

# Calculate Likelihood
  params["llh"] <- EpidemicR::log_likelihood(inf.times, rem.times, rate_mat)
  
# Plot the epidemic
  pdf("../Data/HeteroLatt/epi_plot.pdf") 
  epi_plot = EpidemicR::Plot_epidemic(xy.coords, Hetero_sim, init.inf = c(1))
  while (!is.null(dev.list()))  dev.off() # save plot
```

```{r Save the Hetero Latt epidemic}

epi_list = list(spatial, params, rate_mat, Hetero_sim)

saveRDS(epi_list, file="../Data/HeteroLatt/epi_list.RData")

```





# Heterogeneous Epidemic (Edge spatial distribution)

The case where individuals are distributed in evenly spaced clusters along the left and right edges of the plane, with a large space between the left and right clusters, and individuals within Euclidean distance $d$ of each other make contact at rate $\beta_1$, otherwise at rate $\beta_2$.

First identify a parameter set with an modal total number infected of around 25% of the population.

```{r Identify a parameter set for Heterogeneous (Edge) epidemic}

set.seed(1)

betas = seq(0.001, 0.007, 0.0005)

ps = seq(0.01, 0.1, 0.01)

gammas = seq(0.1, 0.4, 0.02)

ds = seq(0.5, 5, 0.5)

parameter_options = Identify_parameter_set(model = "Edge", N = 100, betas = betas, ps = ps, ds = ds, gammas = gammas, nsims = 1000, xlim = 20, ylim = 20, clusters = 10)

parameter_options

```

In this case we choose $\beta_1 = 0.007$, $\beta_2 = 0.00007$, $d = 5$, and $\gamma = 0.11$.

The following histograms show the distribution of final sizes for this parameter set, and the table provides some seeds that generate an epidemic of the desired size.

```{r Plot histograms for Heterogeneous (Edge) epidemic}

Hist_final_size(model = "Edge", N = 100, betas = c(0.007, 0.00007), d = 5, gamma = 0.11, nsims = 10000, xlim = 20, ylim = 20, clusters = 10)

```

Finally, generate the chosen epidemic and save the results

```{r Generate epidemic 8}

set.seed(719)

# Generate a distance matrix
   spatial <- EpidemicR::Dist_mat_edge(N=100, xlim = 20, ylim = 20, clusters = 10)
   xy.coords <- spatial[[1]]
   distance_mat <- spatial[[2]]

# Generate an associated infection rate matrix
   params <- c("beta1" = 0.007, "beta2" = 0.00007, "gamma" = 0.11, "d" = 5, "p" = 0.01, "llh" = NA)
   rate_mat <- EpidemicR::Beta_mat_form(distance_mat, c(params["beta1"], params["beta2"]), params["d"])
   
# Generate a simulated epidemic
   Hetero_sim <- EpidemicR::GSE_sim(N = 100, beta.mat = rate_mat, gamma = 0.11)
   
# Total number of infected
   n_I = sum(Hetero_sim[,2] < Inf)
  
# Who is infected
   inf.ids <- which(Hetero_sim[,2] < Inf)

# Infection times
  inf.times <- Hetero_sim[,2]

# Removal times
  rem.times <- Hetero_sim[,3]

# Calculate Likelihood
  params["llh"] <- EpidemicR::log_likelihood(inf.times, rem.times, rate_mat)
  
# Plot the epidemic
  pdf("../Data/HeteroEdge/epi_plot.pdf") 
  epi_plot = EpidemicR::Plot_epidemic(xy.coords, Hetero_sim, init.inf = c(1))
  while (!is.null(dev.list()))  dev.off() # save plot
```

```{r Save the Hetero Edge epidemic}

epi_list = list(spatial, params, rate_mat, Hetero_sim)

saveRDS(epi_list, file="../Data/HeteroEdge/epi_list.RData")

```




# Heterogeneous Epidemic (Lines spatial distribution)

The case where individuals are distributed on a lattice where the distance between each successive row and column is increasing, and individuals within Euclidean distance $d$ of each other make contact at rate $\beta_1$, otherwise at rate $\beta_2$.

First identify a parameter set with an modal total number infected of around 25% of the population.

```{r Identify a parameter set for Heterogeneous (Lines) epidemic}

set.seed(1)

betas = seq(0.001, 0.007, 0.0005)

ps = seq(0.01, 0.1, 0.01)

gammas = seq(0.1, 0.4, 0.02)

ds = seq(0.5, 5, 0.5)

parameter_options = Identify_parameter_set(model = "Lines", N = 100, betas = betas, ps = ps, ds = ds, gammas = gammas, nsims = 1000)

parameter_options

```

In this case we choose $\beta_1 = 0.007$, $\beta_2 = 0.00007$, $d = 5$, and $\gamma = 0.11$.

The following histograms show the distribution of final sizes for this parameter set, and the table provides some seeds that generate an epidemic of the desired size.

```{r Plot histograms for Heterogeneous (Unif) epidemic}

Hist_final_size(model = "Lines", N = 100, betas = c(0.007, 0.00007), d = 5, gamma = 0.11, nsims = 10000)

```

Finally, generate the chosen epidemic and save the results

```{r Generate epidemic 9}

set.seed(295)

# Generate a distance matrix
   spatial <- EpidemicR::Dist_mat_lines(N=100)
   xy.coords <- spatial[[1]]
   distance_mat <- spatial[[2]]

# Generate an associated infection rate matrix
   params <- c("beta1" = 0.007, "beta2" = 0.00007, "gamma" = 0.11, "d" = 5, "p" = 0.01, "llh" = NA)
   rate_mat <- EpidemicR::Beta_mat_form(distance_mat, c(params["beta1"], params["beta2"]), params["d"])
   
# Generate a simulated epidemic
   Hetero_sim <- EpidemicR::GSE_sim(N = 100, beta.mat = rate_mat, gamma = 0.11)
   
# Total number of infected
   n_I = sum(Hetero_sim[,2] < Inf)
  
# Who is infected
   inf.ids <- which(Hetero_sim[,2] < Inf)

# Infection times
  inf.times <- Hetero_sim[,2]

# Removal times
  rem.times <- Hetero_sim[,3]

# Calculate Likelihood
  params["llh"] <- EpidemicR::log_likelihood(inf.times, rem.times, rate_mat)
  
# Plot the epidemic
  pdf("../Data/HeteroLines/epi_plot.pdf") 
  epi_plot = EpidemicR::Plot_epidemic(xy.coords, Hetero_sim, init.inf = c(1))
  while (!is.null(dev.list()))  dev.off() # save plot
```

```{r Save the Hetero Edge epidemic}

epi_list = list(spatial, params, rate_mat, Hetero_sim)

saveRDS(epi_list, file="../Data/HeteroLines/epi_list.RData")

```






# Homogeneous Epidemic (Chain Binomial construction)

As a bonus we will generate a discrete time homogeneous epidemic using the Chain-Binomial construction:

```{r Chain Binomial Simulator}

p_exp <- function(I_t, N_t, beta){
  p_exp_t = 1-exp(beta * - (I_t/N_t))
  return(p_exp_t)
}

cbSIRsim <- function(S0, I0, R0, T, beta, gamma){

  data = matrix(data = NA, nrow = 100, ncol = 3)
  track = matrix(data = NA, nrow = 100, ncol = 3)

  St = S0
  It = I0
  Rt = R0
  Nt = St + It + Rt

  p_rem = 1 - exp(-gamma)

  for (t in 1:T){

    p_exp_t = p_exp(It, Nt, beta)

    
    new_I = rbinom(1, St, p_exp_t)
    new_R = rbinom(1, It, p_rem)

    St = St - new_I
    It = It + new_I - new_R
    Rt = Rt + new_R
    Nt = St + It + Rt

    data[t, ] = c(St, It, Rt)
    track[t, ] = c(new_I, new_R, p_exp_t)

  }

  return(list(data, track))
}

```

```{r Identify Seed function}

Hist_final_size_cb <- function(nsims, S0, I0, R0, T, beta, gamma){
  
  n_I = rep(NA, nsims)
  
  valid_seeds = array(NA, nsims)
  
  for(s in 1:nsims){
    
    set.seed(s)
    
    cb_results = cbSIRsim(S0, I0, R0, T, beta, gamma)

    sim = cb_results[[1]]
    events = cb_results[[2]]
    
    n_I[s] = sim[T, 3]
    
    if(n_I[s] == 25){
      valid_seeds[s] = s
    }
  }
  
  print(head(na.omit(as.data.frame(valid_seeds))))

  hist(n_I, breaks = 50)
  abline(v = mean(n_I), # Add line for mean
          col = "red",
          lwd = 3)
  text(x = mean(n_I)+ 11, # Add text for mean
     y = mean(n_I) * nsims * 0.01,
     paste("Mean =", mean(n_I)),
     col = "red",
     cex = 1)
  abline(v = median(n_I),# Add line for median
         col = "blue",
         lwd = 3)
  text(x = median(n_I)+ 9, # Add text for median
     y = median(n_I) * nsims * 0.1,
     paste("Median =", median(n_I)),
     col = "blue",
     cex = 1)
}

```

```{r Identify Seed}

S0 = 99
I0 = 1
R0 = 0
T = 100
beta = 0.11
gamma = 0.10

Hist_final_size_cb(100000, S0, I0, R0, T, beta, gamma)

```



```{r Generate Simulated Data}

S0 = 99
I0 = 1
R0 = 0
T = 100
beta = 0.11
gamma = 0.10

set.seed(314)

cb_res = cbSIRsim(S0, I0, R0, T, beta, gamma)

cb_data = cb_res[[1]]
cb_track = cb_res[[2]]

plot(1:100, cb_data[,1], type = "l", col = "darkgreen", ylim = c(0, 100))
  lines(1:100, cb_data[,2], type = "l", col = "red")
  lines(1:100, cb_data[,3], type = "l", col = "blue")

```

```{r Save the Hetero Edge epidemic}

epi_list = list(c(beta, gamma), cb_data, cb_track)

saveRDS(epi_list, file="../Data/cbHomo/epi_list.RData")

```












