---
output:
  pdf_document: default
  html_document: default
title: "Unpaired Homogeneous General Stochastic Epidemic Inference"
author: "Benjamen Simon"
date: "September 2022"
knit: (function(inputFile, encoding) { 
          rmarkdown::render(
            inputFile,
            encoding = encoding, 
            output_dir = "../Reports") 
            })
header-includes:
  - \usepackage{amsmath}
  - \usepackage{enumerate}
  - \usepackage{bm}
  - \usepackage[T1]{fontenc}
---

```{r setup, include=FALSE}

# Rmarkdown settings
knitr::opts_chunk$set(echo = FALSE, cache = TRUE, eval= TRUE, message = FALSE, cache.path = "Caches/", fig.path = "Figures/Homo/")

# Default to non-scientific notation
options(scipen=999)

# Suppress summarise info
options(dplyr.summarise.inform = FALSE)

```

```{r Packages}

  ## Data manipulation
  library(tidyverse)

  ## Plotting
  library(ggplot2)
  library(ggcorrplot)
  source("VisualisationFunctions.R")
  library(plotMCMC) # autocorrelation plots and other diagnostic plots

  ## MCMC Diagnostics
  library(coda)

  ## Epidemics
  library(EpidemicR) # For simulating and making inference on GSE epidemics

```

# Load in the epidemic

The epidemic we have simulated has a modal number of infected individuals as roughly 25% of the population

```{r Load epidemic}

epi_list <- readRDS(file="../Data/Homo/epi_list.RData")

# Distance matrix
  spatial <- epi_list[[1]]
  xy.coords <- spatial[[1]]
  distance_mat <- spatial[[2]]

# Associated infection rate matrix
  params <- epi_list[[2]]
  rate_mat <- epi_list[[3]]
   
# Simulated epidemic
  Homo_sim <- epi_list[[4]]

# Total number of infected
  n_I = sum(Homo_sim[,2] < Inf)
  
# Who is infected
  inf.ids <- which(Homo_sim[,2] < Inf)

# Infection times
  inf.times <- Homo_sim[,2]

# Removal times
  rem.times <- Homo_sim[,3]
  
# Plot the epidemic
  epi_plot = EpidemicR::Plot_epidemic(xy.coords, Homo_sim, init.inf = c(1))

```

# Component functions of the MCMC algorithm

## The likelihood / posterior

```{r Calculte integrals}

int.calc <- function(R = rem.times, I = inf.times, N){
  
  I <- I[which( I < Inf )]
  R <- R[which( R < Inf )]
  
  I <- sort(I)
  R <- sort(R)
  
  n.I = length(I)
  
  # Calculate the double sum for the infection part
  
  sum_n_n <- matrix(nrow = n.I, ncol = n.I)
  
  for(i in 1:n.I){
    for(j in 1:n.I){
      sum_n_n[i,j] <- (  min( c(R[i], I[j]) ) - min( c(I[i], I[j]) )  )
    }
  }
  
  sum_n_n <- sum(sum_n_n)
  
  # Calculate the single sum used in both parts
  
  sum_n <- sum(R-I)
  
  # Evaluate the infection and removal integrals
  
  int.xy <- sum_n_n + (N-n.I) * sum_n
  int.y <- sum_n
  
  return(c(int.xy, int.y))
}

```

```{r Calculate product}

log.prod.calc <- function(R = rem.times, I = inf.times, N){
  
  inf.ids <- which(I < Inf)
  
  kappa <- which.min(I[inf.ids])
  
  n.I = length(inf.ids)
  
  num.suscept <- rep(NA, (n.I - 1))
  num.infect <- rep(NA, (n.I - 1))
  
  j=1
  
  for(i in inf.ids[-kappa]){
    
    num.suscept[j] <- sum(I >= I[i]) # >= because looking at just before
    
    num.infect[j] <-  N - num.suscept[j] - sum(R < I[i])
    
    j = j+1
  }
  
  log.prod <- sum(log( num.suscept * num.infect ))

  return(log.prod)
}
  
```

Thus we can now bring the two parts together and calculate the log-likelihood (more accurately the log-posterior for the infection times);

```{r Calculate likelihood}

simple_llh <- function(inf.times, rem.times, N, beta, delta){
  
  ints <- int.calc(R = rem.times, I = inf.times, N)
    
  int.xy <- ints[1]
  int.y <- ints[2]
  
  prod <- log.prod.calc(R = rem.times, I = inf.times, N)
    
  llh <- prod - beta * int.xy - delta * int.y
  
  return(llh)
}

```

## Metropolis-Hastings Acceptance Probability

```{r Calculate MH accept prob}

alpha.calc <- function(llh, llh.prime){
	alpha <- exp(llh.prime - llh)
	return(min(c(1, alpha)))
}

```

## The MCMC algorithm

Assuming we know the true infection times, we can remove the Metropolis-Hastings step and reduce the algorithm down to just a Gibbs sampler.

```{r Simple MCMC algorithm with I known}

Simple_MCMC_Iknown <- function(N, n.I, inf.times, rem.times, N.its, infupdate, lambda.b, nu.b, lambda.g, nu.g){
  
  ##############################################
  ### Initialise the parameters and epidemic ###
  ##############################################
    
      rem.times <- sort(rem.times)
      inf.times <- sort(inf.times)
      n.I <- sum(rem.times < Inf)

  ######################    
  ### Results matrix ###
  ######################
      
      res <- matrix(ncol = 4, nrow = N.its)
      colnames(res) <- c("sample", "beta", "delta", "llh")
  
  ##########################
  ### Functional objects ###
  ##########################
      
      it = 1
      
      ints <-  int.calc(R = rem.times, I = inf.times, N)
      
      acc.sum.I = 0
  
  ###########################
  ### ~~ THE ALGORITHM ~~ ###
  ###########################
     
      while(it <= N.its){

            ##############################
            ### Gibbs sampler for beta ###
            ##############################
        
                beta.cur <- rgamma(n=1, shape = (n.I-1+nu.b), rate = (lambda.b + ints[1]))
                
            ###############################
            ### Gibbs sampler for delta ###
            ###############################
        
                delta.cur <- rgamma(n=1, shape = (n.I+nu.g), rate = (lambda.g + ints[2]))
                
                llh <- simple_llh(inf.times, rem.times, N, beta.cur, delta.cur)
                
                # Record beta and delta
                res[it,] <- c(it, beta.cur, delta.cur, llh)
                
                # Update count
                it = it + 1
      }
  
  return(list(res, (acc.sum.I/N.its)))
}

```

As is more often the case, when the infection times are unknown we use MH methods

```{r Simple MCMC algorithm}

Simple_MCMC <- function(N, n.I, rem.times, N.its, infupdate, lambda.b, nu.b, lambda.g, nu.g){
  
  ##############################################
  ### Initialise the parameters and epidemic ###
  ##############################################
    
      rem.times <- sort(rem.times)
      n.I <- sum(rem.times < Inf)
      
      repeat{
        
        beta.cur <- runif(1, 0, 0.1)
        delta.cur <- runif(1, 0, 0.2)
        
        inf.times <- c(0, rep(Inf, (N-1)))
        
    		for(i in 2:n.I){
          inf.times[i] <- runif(1, 0, rem.times[(i-1)])
        }
        inf.times <- sort(inf.times)
    
    		if(simple_llh(inf.times, rem.times, N, beta.cur, delta.cur) != -Inf){ break }
      }
      
      print("initialised")
  
  ##########################################
  ### Start and end time of the epidemic ###
  ##########################################
      
      inf.ids <- which(rem.times < Inf)
      kappa <- which.min(inf.times[inf.ids])
      
      I1 <- inf.times[which.min(inf.times)]
      T <- max(rem.times[inf.ids])
      
  ######################    
  ### Results matrix ###
  ######################
      
      res <- matrix(ncol = 4, nrow = N.its)
      colnames(res) <- c("sample", "beta", "delta", "llh")
  
  ##########################
  ### Functional objects ###
  ##########################
      
      it = 1
      
      llh <- simple_llh(inf.times, rem.times, N, beta.cur, delta.cur)
      ints <-  int.calc(R = rem.times, I = inf.times, N)
      
      acc.sum.I = 0
  
  ###########################
  ### ~~ THE ALGORITHM ~~ ###
  ###########################
     
      while(it <= N.its){

            ##############################
            ### Gibbs sampler for beta ###
            ##############################
        
                beta.cur <- rgamma(n=1, shape = (n.I-1+nu.b), rate = (lambda.b + ints[1]))
                
            ###############################
            ### Gibbs sampler for delta ###
            ###############################
        
                delta.cur <- rgamma(n=1, shape = (n.I+nu.g), rate = (lambda.g + ints[2]))
                
                llh <- simple_llh(inf.times, rem.times, N, beta.cur, delta.cur)
                      
            ###################################
            ### MH step for infection times ###
            ###################################

                # Which infection time is being replaced
                kappa <- which.min(inf.times[inf.ids])
                Ireplace <- sample(inf.ids[-kappa], infupdate)

                # Draw new infection times
                I1 <- inf.times[1]
                Idraw <- runif(infupdate, I1, T)

                inf.times.prime <- inf.times
                inf.times.prime[Ireplace] <- Idraw
                inf.times.prime <- sort(inf.times.prime)

                # Calculate functional objects for new infection times

                llh.prime <- simple_llh(inf.times.prime, rem.times, N, beta.cur, delta.cur)

                # MH acceptance probability
                alpha <- alpha.calc(llh, llh.prime)
                
                # Do we accept the new time or not?
                accept.test <- runif(1,0,1)
                if(accept.test < alpha){ #If yes:
                  
                  inf.times <- inf.times.prime
                  
                  ints <- int.calc(R = rem.times, I = inf.times, N)
                  llh <- llh.prime
                  
                  acc.sum.I = acc.sum.I+1
                }
                
                # Record beta and delta
                res[it,] <- c(it, beta.cur, delta.cur, llh)
                
                # Update count
                it = it + 1
      }
  
  return(list(res, (acc.sum.I/N.its)))
}

```

# Making inference (Infection times known)

```{r Performing analysis I known}

set.seed(412)

## Run the algorithm

start.time <- Sys.time()

res.list.Iknown <- Simple_MCMC_Iknown(N=100, n.I=n_I, inf.times, rem.times, N.its = 100000, infupdate = 1, lambda.b = 1, nu.b = 1, lambda.g = 1, nu.g = 1)

end.time <- Sys.time()

res.Iknown <- res.list.Iknown[[1]]


## Summarise the results

summary(res.Iknown[-(1:5000),])

```

```{r}

hex_beta_delta_I_known = gg_hex_plot_homo(res.Iknown[-c(1:10000), ], params_true, c(0.006,0.7,  0.006,0.6,  0.006,0.5,  0.006,0.4), c(0, 0.008), c(0, 0.8))

hex_beta_delta_I_known

ggsave(filename = "../Visualisation/Homo/hex_beta_delta_I_known.png", plot = hex_beta_delta_I_known, width = 10, height = 8)

```



```{r Plot the results I known}

## Plot the results

trace_plot_I_known = gg_trace_plot_homo(res.Iknown, params, 10000, c(20000, 0.0005, 20000, 0.03), c(0, 0.004, 0, 0.4))

hist_plot_I_known = gg_hist_plot_homo(res.Iknown, params, 10000, c(0.004, 700, 0.3, 8), c(0, 0.005, 0, 0.35))

hist_plot_w_prior_I_known = gg_hist_plot_w_prior_homo(res.Iknown, params, 10000, c(1, 1, 1, 1), c(0.004, 700, 0.3, 8), c(0, 0.005, 0, 0.35))


## Save the results

ggsave(filename = "../Visualisation/Homo/trace_I_known.png", plot = trace_plot_I_known, width = 10, height = 8)

ggsave(filename = "../Visualisation/Homo/hist_I_known.png", plot = hist_plot_I_known, width = 10, height = 8)

ggsave(filename = "../Visualisation/Homo/hist_w_prior_I_known.png", plot = hist_plot_w_prior_I_known, width = 10, height = 8)

```

The MCMC algorithm for all parameters took `r round(end.time - start.time, 2)` minutes to run.

The algorithm estimates the mean, median, and quartiles of $\beta$ to be `r mean(res.Iknown[-(1:5000),2])` (`r quantile(res.Iknown[-(1:5000),2], probs = c(.25, .5, .75))`) compared to the true value of $\beta =$  `r params[1]`, with an effective sample size of `r round(coda::effectiveSize(res.Iknown[,2]))`. It estimates the mean, median, and quartiles of $\delta$ to be `r mean(res.Iknown[-(1:5000),3])` (`r quantile(res.Iknown[-(1:5000),3], probs = c(.25, .5, .75))`) compared to the true values of $\delta =$ `r params[3]`, with an effective sample size of `r round(coda::effectiveSize(res.Iknown[,3]))`.


```{r}

png("../Visualisation/Homo/autocorr_Iknown.png", width=1000, height=800)
plotAuto(res.Iknown[, c(2,3)], lag.max=50)
dev.off()

```


# Making inference (Infection times unknown)

Using our simulated epidemic from earlier, we will now apply our MCMC algorithm in an effort to make inference and recover the parameters.

```{r Performing analysis}

set.seed(403)

## Run the algorithm

start.time <- Sys.time()

res.list <- Simple_MCMC(N=100, n.I=n_I, rem.times, N.its = 100000, infupdate = 1, lambda.b = 1, nu.b = 1, lambda.g = 1, nu.g = 1)

end.time <- Sys.time()

res <- res.list[[1]]

## Summarise the results

print(c("acceptance probability = ", res.list[[2]]))

summary(res[-(1:5000),])

```



```{r}

hex_beta_delta_I_unknown = gg_hex_plot_homo(res[-c(1:10000), ], params_true, c(0.001,0.8,  0.001,0.73,  0.001,0.66,  0.001,0.59), c(0, 0.008), c(0, 0.8))

hex_beta_delta_I_unknown

ggsave(filename = "../Visualisation/Homo/hex_beta_delta_I_unknown.png", plot = hex_beta_delta_I_unknown, width = 10, height = 8)

```


```{r}

contour_beta_delta_I_unknown = gg_contour_plot_homo(res, params, 10000,  c(0, 0.008), c(0.1, 0.5))

contour_beta_delta_I_unknown

ggsave(filename = "../Visualisation/Homo/contour_beta_delta_I_unknown.png", plot = contour_beta_delta_I_unknown, width = 10, height = 8)

```


```{r Plot the results}

## Plot the results

trace_plot_I_unknown = gg_trace_plot_homo(res, params, 10000, c(20000, 0.0005, 20000, 0.05), c(0, 0.007, 0, 0.7))

hist_plot_I_unknown = gg_hist_plot_homo(res, params, 10000, c(0.006, 350, 0.6, 4), c(0, 0.007, 0, 0.7))

hist_plot_w_prior_I_unknown = gg_hist_plot_w_prior_homo(res, params, 10000, c(1, 1, 1, 1), c(0.006, 350, 0.6, 4), c(0, 0.007, 0, 0.7))


## Save the results

ggsave(filename = "../Visualisation/Homo/trace_I_unknown.png", plot = trace_plot_I_unknown, width = 10, height = 8)

ggsave(filename = "../Visualisation/Homo/hist_I_unknown.png", plot = hist_plot_I_unknown, width = 10, height = 8)

ggsave(filename = "../Visualisation/Homo/hist_w_prior_I_unknown.png", plot = hist_plot_w_prior_I_unknown, width = 10, height = 8)

```

The MCMC algorithm for all parameters took `r round(end.time - start.time, 2)` minutes to run.

The algorithm estimates the mean, median, and quartiles of $\beta$ to be `r mean(res[-(1:5000),2])` (`r quantile(res[-(1:5000),2], probs = c(.25, .5, .75))`) compared to the true value of $\beta =$ `r params[1]`, with an effective sample size of `r round(coda::effectiveSize(res[,2]))`. It estimates the mean, median, and quartiles of $\delta$ to be `r mean(res[-(1:5000),3])` (`r quantile(res[-(1:5000),3], probs = c(.25, .5, .75))`) compared to the true values of $\delta =$ `r params[3]`, with an effective sample size of `r round(coda::effectiveSize(res[,3]))`.


```{r}

png("../Visualisation/Homo/autocorr_Iunknown.png", width=1000, height=800)
plotAuto(res[, c(2,3)], lag.max=50)
dev.off()

```
